---
title: "partitioning_AS_AN_new"
output: html_document
---

Goal is to look at the strength of the storage effect in seedbanking annual (AS) and nonseedbanking annual (AN) species interactions. We want to measure how the storage effect [and the relative strength of storage effect:relative nonlinearity] changes with 1. magnitude of envrionmental variation 2. autocorrelation of environmental variation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

## Personal ggplot theme
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
    #  strip.text=element_text(size=12),
    #  axis.text=element_text(size=12),
      complete = TRUE
    )
}
```

## Parameters and set up
Setting parameters we're going to use for alpha and lambda under different environmental conditions. Starting with the lambda and alpha parameters for an (as our AN) and as (as our AS) from Hallett et al. 2019 Ecology Letters paper, using the "wet" treatment as our wet environment and the "dry" treatment for our dry environment. Also going to use the seedbank parameters for as listed in that ms, which references Rice 1985, and simplifying the an to being non-seedbanking (so shifting its survival in the seed bank proportion from 0.4 to 0)
*This is different from the annual-perennial parameters*
```{r parameters}

# seed survival and germination fractions
# don't currently differ by environmental conditions but they could
as <- 0 # changed to 0 to be a fully non-seedbanking species
ag <- 0.9
es <- 0.82
eg <- 0.6

# main parameters
pars <- read.csv('model_parameters_hallett.csv') %>%
  mutate(species = ifelse(species == 'Avena', 'an', 'as'), # generalize species names
         treatment = ifelse(treatment == 'consistentDry','dry','wet')) %>%
  rename(a.an = aiA, a.as = aiE) %>%
  select(-X) %>%
  mutate(seed.surv = ifelse(species == 'an', as, es),
         seed.germ = ifelse(species == 'an', ag, eg)) # add seed survival and germ columns

# adding average values to the parameter df
an.means <- pars %>%
  filter(species == 'an') %>% 
  select(a.an, a.as, lambda, seed.surv, seed.germ) %>%
  colMeans()
as.means <- pars %>%
  filter(species == 'as') %>% 
  select(a.an, a.as, lambda, seed.surv, seed.germ) %>%
  colMeans()
pars <- pars %>%
  add_row(treatment = 'average', species = 'an', 
          a.an = an.means['a.an'], a.as = an.means['a.as'], lambda = an.means['lambda'],
          seed.surv = an.means['seed.surv'], seed.germ = an.means['seed.germ']) %>%
  add_row(treatment = 'average', species = 'as', 
          a.an = as.means['a.an'], a.as = as.means['a.as'], lambda = as.means['lambda'],
          seed.surv = as.means['seed.surv'], seed.germ = as.means['seed.germ'])
```

Functions for one step of species interaction calculations. Trying to make it general to whether it's both species present or just one--if it's one, that population is just at 0. Takes in the full parameter df for all the species (just for one environmental condition though) and a list of species population sizes. If it's just one species running to equilibrium then it's N0 will be 0. 
*this is different from the annual-perennial interactions*
```{r species-interactions}
pop_interactions <- function(N0, param){
  pars.as <- param %>% filter(species == 'as')
  pars.an <- param %>% filter(species == 'an')
  
  n.as <- N0$as * (1 - pars.as$seed.germ) * pars.as$seed.surv + 
    N0$as * pars.as$seed.germ *pars.as$lambda /
    (1 + pars.as$a.an * pars.an$seed.germ * N0$an +
       pars.as$a.as * pars.as$seed.germ * N0$as)
  
  n.an <- N0$an * pars.an$seed.germ *pars.an$lambda /
    (1 + pars.an$a.an * pars.an$seed.germ * N0$an +
       pars.an$a.as * pars.as$seed.germ * N0$as)
  
  n.all <- tibble(an = n.an,
                  as = n.as)
  return(n.all)
}

```

## Running models
*only thing that changes is the tibble column names in the first step and the lambda columns change from c('lambda','seed.germ','stem.surv') to c('lambda','seed.germ','seed.surv'). This second part could be easily generalized by de-selecting treatment, species, and alpha columns instead*
Functions in original script (partioning_P_AN_new.Rmd) is that sp.invader and sp.resident parameters can be either a single value (ex: 'a') or multiple values (ex: c('s','p')). This is adjusted for 'an' and 'as' only -- didn't take out the steps of finding the stage distribution for the perennial species since it should automatically skip that for an invader with only one stage. 

*For partitions where lambda is set to not vary, also setting germination and survival terms to not vary*
```{r running-invasion}
run_invasion <- function(sp.invader, sp.resident, N0.r, parameters, env.cond,
                         vary.lambda = TRUE, vary.alpha = TRUE, covary = TRUE,
                         init.time.steps = 50, total.time.steps = 100) {
  
  
  # populations with resident species not impacted by the invader
  # NOTE: this isn't generic to any invader and resident species names
  n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                  as = as.numeric(rep(NA, total.time.steps)))
  
  # setting initial populations
  n.all[1,sp.resident] <- N0.r
  n.all[1,sp.invader] <- 0
  
  # name of column(s) for resident species pop when impacted by invading pop
  sp.res.impact <- paste(sp.resident, 'impact', sep = '.')
  
  
  # parameters for the average env condition
  p.ave <- parameters %>% filter(treatment == 'average')
  
  # environmental condition vectors for the no covariance option
  env.lambda <- sample(env.cond, size = length(env.cond), replace = FALSE)
  env.alpha <- sample(env.cond, size = length(env.cond), replace = FALSE)
  
  for(t in 1:init.time.steps){
    # parameters for the varying env condition
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
    # adjusting non-varying or non-covarying parameters
    lambda.columns <- c('lambda','seed.germ','seed.surv') # NOTE would need to change for a generalized function with different parameter columns
    alpha.columns <- c(paste('a', sp.invader, sep = '.'), 
                       paste('a', sp.resident, sep = '.'))
    
    # constant lambda
    if(vary.lambda == FALSE){
      p.env[,lambda.columns] <- p.ave[,lambda.columns]
    }
    
    # constant alpha
    if(vary.alpha == FALSE){
      p.env[,alpha.columns] <- p.ave[,alpha.columns]
    }
    
    # parameters for no covariance between alpha and lambda
    if(covary == FALSE){
      lambda.env  <- parameters %>% filter(treatment == env.lambda[t])
      alpha.env <- parameters %>% filter(treatment == env.alpha[t])
      p.env[,lambda.columns] <- lambda.env[,lambda.columns]
      p.env[,alpha.columns] <- alpha.env[,alpha.columns]
    }
    
    # running one time step with the appropriate environmental parameters
    n.all[t+1,] <- pop_interactions(N0 = n.all[t,], param = p.env)
    
  }
  
  # figuring out rough invader population distribution if the invader is perennial
  # using average parameters and resident population size during the final warm-up step
  # makes it faster for running in for loop
  if(length(sp.invader) > 1) {
    n.inv.start <- 1/length(sp.invader)
    n.old <- n.all[init.time.steps,]
    n.old[sp.invader] <- n.inv.start
    n.old.inv <- n.inv.start[1]
    change <- 1
    
    while(change > 0.1){
      n.update <- pop_interactions(N0 = n.old, param = p.ave)
      n.per.total <- sum(n.update[sp.invader])
      n.update[sp.invader] <- n.update[sp.invader]/n.per.total
      n.update[sp.resident] <- n.old[sp.resident]
      n.old <- n.update
      
      n.new.inv <- n.update[sp.invader]
      change <- abs(n.old.inv - n.new.inv[1])
      n.old.inv <- n.new.inv[1]
    }
  } 
  
  # running the resident and invader for the remaining time steps
  for (t in (init.time.steps+1):(total.time.steps-1)){
    # parameters for the varying env condition
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
   # adjusting non-varying or non-covarying parameters
    lambda.columns <- c('lambda','seed.germ','seed.surv') # NOTE would need to change for a generalized function with different parameter columns
    alpha.columns <- c(paste('a', sp.invader, sep = '.'), 
                       paste('a', sp.resident, sep = '.'))
    
    # constant lambda
    if(vary.lambda == FALSE){
      p.env[,lambda.columns] <- p.ave[,lambda.columns]
    }
    
    # constant alpha
    if(vary.alpha == FALSE){
      p.env[,alpha.columns] <- p.ave[,alpha.columns]
    }
    
    # parameters for no covariance between alpha and lambda
    if(covary == FALSE){
      lambda.env  <- parameters %>% filter(treatment == env.lambda[t])
      alpha.env <- parameters %>% filter(treatment == env.alpha[t])
      p.env[,lambda.columns] <- lambda.env[,lambda.columns]
      p.env[,alpha.columns] <- alpha.env[,alpha.columns]
    }
    
    
    # running invader and resident together
    N0.together <- n.all[t,]
    
    # figuring out starting invader population distribution if the invader is perennial
    # starting from the n.old found for average conditions above
    if(length(sp.invader) > 1){
      change <- 1
      while(change > 0.01){
        n.update <- pop_interactions(N0 = n.old, param = p.env)
        n.per.total <- sum(n.update[sp.invader])
        n.update[sp.invader] <- n.update[sp.invader]/n.per.total
        n.update[sp.resident] <- n.old[sp.resident]
        n.old <- n.update
        n.new.inv <- n.update[sp.invader]
        change <- abs(n.old.inv - n.new.inv[1])
        n.old.inv <- n.new.inv[1]
      }
      N0.together[sp.invader] <- n.update[sp.invader] # using the equilibrium stage distribution from above
    } else {
      N0.together[sp.invader] <- 1 # if the invader has only one life stage
    }
    
    
    # running one time step with the appropriate environmental parameters
    n.res.impact <- pop_interactions(N0 = N0.together, param = p.env)
    
    # saving this output as the invader population and the impacted resident population
    n.all[t+1,sp.invader] <- n.res.impact[sp.invader]
    n.all[t+1,sp.res.impact] <- n.res.impact[sp.resident]
    
    # a version of this with the resident unaffected by the invader
    N0.res.only <- n.all[t,] %>% select(all_of(sp.invader), all_of(sp.resident))
    N0.res.only[sp.invader] <- 0
    n.res.only <- pop_interactions(N0 = N0.res.only, param = p.env)
    n.all[t+1,sp.resident] <- n.res.only[sp.resident]
    
    # growth rate of the resident
    sp.res.gr <- paste('gr', paste(sp.resident, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.res.gr] <- sum(n.all[t+1,sp.res.impact])/sum(n.all[t,sp.resident])
    
    # growth rate of the invader
    sp.inv.gr <- paste('gr', paste(sp.invader, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.inv.gr] <- sum(n.all[t+1,sp.invader])
    
  }
  
  n.all$time <- 1:total.time.steps
  return(n.all)
  
}

```

Plotting function so we can visualize what's going on with the different types of variation *unchanged*
```{r plotting-function}
plot_invasion <- function(df){ # plots actual population sizes
  df.long <- df %>% 
    pivot_longer(cols = !c(time, starts_with('gr')),
                 names_to = 'species', values_to = 'n',
                 values_drop_na = TRUE) %>% 
    mutate(n.log = log(n))

  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    theme_cw() +
    xlab('Time') + 
    ylab('Population size (log)')
  
}

plot_gr <- function(df) { # plots population growth rates
  df.long <- df %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', values_to = 'n',
                 values_drop_na = TRUE) %>% 
    mutate(n.log = log(n))
  
  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    geom_point() +
    theme_cw() +
    xlab('Time') + 
    ylab('Growth Rate (log)')
}
```

# Running and plotting invasions 
*unchanged*
Full variation (normal run)
```{r normal-run}
time.warm.up <- 200
time.full <- 300

# two equally probable environmental conditions
env.draw <- rbinom(time.full, 1, 0.5) 
env.condition <- if_else(env.draw > 0, 'wet','dry')
  
invade.as.full <-  run_invasion(sp.invader = 'as', sp.resident = 'an', N0.r = 100, 
                         parameters = pars, env.cond = env.condition,
                         vary.lambda = TRUE, vary.alpha = TRUE, 
                         covary = TRUE,
                         init.time.steps = time.warm.up, total.time.steps = time.full)

plot_invasion(invade.as.full)
plot_gr(invade.as.full)



invade.an.full <-  run_invasion(sp.invader = 'an', sp.resident = 'as', N0.r = 100, 
                         parameters = pars, env.cond = env.condition,
                         vary.lambda = TRUE, vary.alpha = TRUE, 
                         covary = TRUE,
                         init.time.steps = time.warm.up, total.time.steps = time.full)

plot_invasion(invade.an.full)
plot_gr(invade.an.full)

```

# Calculating partitions 
*unchanged*
Function to calculate partitions for a given invader and resident
*Could add another small function so we don't have to copy and paste the filter, pivot, mutate, select functions for each step?*
```{r partition-function}
partition_epsilons <- function(sp.inv, sp.res){
  
  time.warm.up <- 200
  time.full <- 300
  
  # two equally probable environmental conditions
  env.draw <- rbinom(time.full, 1, 0.5) 
  env.condition <- if_else(env.draw > 0, 'wet','dry')
  
  res.name <- paste(sp.res, collapse = '.')
  inv.name <- paste(sp.inv, collapse = '.')
  
  part <- tibble(sp.invader = inv.name, sp.resident = res.name)
  
  # no variation
  invade.none <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = FALSE, vary.alpha = FALSE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  part$inv.e0 <- invade.none %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  part$res.e0 <- invade.none %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  # lambda variation only
  invade.l <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                            N0.r = 100, parameters = pars, env.cond = env.condition,
                            init.time.steps = time.warm.up, total.time.steps = time.full,
                            vary.lambda = TRUE, vary.alpha = FALSE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  l.inv.mean <- invade.l %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  l.res.mean <- invade.l %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.el <- l.inv.mean - part$inv.e0 
  part$res.el <- l.res.mean- part$inv.e0
  
  # alpha variation only
  invade.a <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                            N0.r = 100, parameters = pars, env.cond = env.condition,
                            init.time.steps = time.warm.up, total.time.steps = time.full,
                            vary.lambda = FALSE, vary.alpha = TRUE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  a.inv.mean <- invade.a %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  a.res.mean <- invade.a %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.ea <- a.inv.mean - part$inv.e0 
  part$res.ea <- a.res.mean- part$inv.e0
  
  # full variation
  invade.full <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = TRUE, vary.alpha = TRUE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  part$inv.full <- invade.full %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  part$res.full <- invade.full %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.eint <- part$inv.full - (part$inv.e0 + part$inv.el + part$inv.ea)
  part$res.eint <- part$res.full - (part$res.e0 + part$res.el + part$res.ea)
  
  # # no correlation
   invade.nocov <- run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = TRUE, vary.alpha = TRUE,
                               covary = FALSE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
   
   part$inv.nocov <- invade.nocov %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
   part$res.nocov <- invade.nocov %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
   # storage effect
   part$inv.storage <- part$inv.eint - part$inv.nocov
   part$res.storage <- part$res.eint - part$res.nocov
  
  return(part)
}


```

Running that partition function 
*this changes with the species codes in the call to the partition_epsilons() function and the species_names() labeller*
```{r partition-run}
time.warm.up <- 200
time.full <- 300

partitions <- partition_epsilons(sp.inv = 'as', sp.res = 'an') %>%
  rbind(partition_epsilons(sp.inv = 'an', sp.res = 'as')) %>%
  #pivot_longer(cols = inv.e0:res.eint) %>%
  pivot_longer(cols = inv.e0:res.storage) %>%
  separate(name, into = c('player', 'partition')) %>%
  pivot_wider(names_from = player, values_from = value) %>%
  mutate(delta = inv - res,
         partition = factor(partition, 
                            levels = c('full','e0','ea','el','eint',
                                       'storage','nocov')))

# This is not currently generalized
species_names <- c(
  an = 'AN invades AS',
  as = 'AS invades AN'
)

colors <- c("#5445b1", "#749dae", "#f3c483", "#5c1a33", "#cd3341", "#f7dc6a", "#4DA896")
colors2 <- c("#749dae", rep('darkslategrey', times = 4), rep("#f3c483", times = 2))

ggplot(partitions, aes(x = partition, y = delta, fill = partition)) + 
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_cw() +
  ylab('Partitioning of growth rate when rare') +
  scale_x_discrete(labels = c('full' = 'Full',  'e0' = expression(Delta^0), 'ea' = expression(Delta^alpha),
                              'el' = expression(Delta^lambda), 'eint' = expression(Delta^{alpha*lambda}),
                              'storage' = expression(Delta^(alpha*lambda)), 'nocov' = expression(Delta^{alpha!=lambda})),
                   name = 'Partition') +
  scale_fill_manual(values = colors2) +
  guides(fill = 'none')

 #ggsave('partitions_AS_AN.pdf', width = 8, height = 5, units = 'in')

```