---
title: "Simulated Stipa Avena data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstan)
library(tidybayes)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

## Simulate data
This is a SIMULATED data set to start thinking about how our models will work with the experimental data we're planning on collecting, and how best to proceed with both the experiment and the analysis. *This is NOT real data!* 

Plots
```{r plots}
background <- c('ss.full', 'ss.half', 'sp.full', 'sp.half', 'av.full', 'av.half','none')
water <- c('dry','wet')
replicates <- 1:8
focal <- c('ss','sp','av')

data.sim <- expand_grid(replicates, water, background, focal)

```

Set up initial parameters, starting with mean parameters/ranges from Mordecai et al 2015, Appendix A. Note that some of these (especially the competition coefficients) are REALLY broad intervals (and seem to be from multiple plant species rather than just stipa - avena). Let's assume those are the parameters for the wet condition, and then add parameters for the dry condition. Lower survival and fecundity for both species, but avena taking a bigger loss. Not changing the competition parameters for now. 

```{r parameters}
species <- c('av','ss','sp')
parameters <- expand_grid(species, water)

g.av <- 0.45 # 0.22 - 0.67 highest posterior density interval
g.ss <- 0.34 # 0.14 - 0.54
parameters$germ <- c(g.av, g.av, g.ss, g.ss, NA, NA)

s.ss.wet <- 0.28 # 0.18 - 0.41
s.ss.dry <- s.ss.wet * 0.5
s.sp.wet <- 0.88 # 0.84 - 0.93
s.sp.dry <- s.sp.wet * 0.8

parameters$surv <- c(0, 0, s.ss.dry, s.ss.wet, s.sp.dry, s.sp.wet)

lambda.av.wet <- 29 # 23 - 37 avena seed production
lambda.av.dry <- lambda.av.wet * 0.5
lambda.sp.wet <- 201 # 173 - 229 stipa seed production
lambda.sp.dry <- lambda.sp.wet * 0.7

parameters$lambda <- c(lambda.av.dry, lambda.av.wet, NA, NA, lambda.sp.dry, lambda.sp.wet)

a.av.av <- 0.00174 # 0.00028 - 0.0031 intraspecific 
a.av.sp <- 0.024 # 0.00001 - 0.067 effect of stipa adults on avena (seed production)
a.av.ss <- a.av.av * 0.5 # multiplier of 0.02 - 0.97 effect of stipa seedlings on avena

a.sp.av <-  0.0138 # 0.00002 - 0.036 effect of avena on stipa adults (seed production)
a.sp.sp <- 0.15 # 0.00025 - 0.34
a.sp.ss <- a.sp.av * 0.5 # multiplier of 0.02 - 0.97 effect of stipa seedlings on stipa adults

a.ss.av <- 0.00284 # 0.000002 - 0.0081 effect of avena on stipa seedlings (survival) 
a.ss.ss <- 0.087 # 0.00003 - 0.24 I'm surprised this is much larger than avena effects above
a.ss.sp <- 0.12 # 0.00005 - 0.36 effect of stipa adults on stipa seedlings

# effect of column j species on row i species
parameters$a.av <- c(a.av.av, a.av.av, a.ss.av, a.ss.av, a.sp.av, a.sp.av) # avena effects on everyone
parameters$a.ss <- c(a.av.ss, a.av.ss, a.ss.ss, a.ss.ss, a.sp.ss, a.sp.ss) # stipa seedling effects on everyone
parameters$a.sp <- c(a.av.sp, a.av.sp, a.ss.sp, a.ss.sp, a.sp.sp, a.sp.sp) # stipa adult effects on everyone

```


Actually I'm not sure we want to use these updated ones or just use the main parameter dataset and add noise as we go, since we also need the germination and survival data for the other species too. This currently works, but doesn't have any noise added in at all (since adding noise at the end was making things negative and I want to think about how best to do it). Now that I have the map functions figured out, I could do it with the dataframe above and the pmap_dfr() approach? Also we want growth rates/some values for the resident communities
```{r calculating seeds out}

background.density <- tibble(background = unique(data.sim$background))
background.density$av <- c(0, 0, 0, 0, 1600, 400, 0)
background.density$ss <- c(1600, 400, 0, 0, 0, 0, 0)
background.density$sp <- c(0, 0, 10, 5, 0, 0, 0)

focal.density <- tibble(focal = c('av','ss','sp'), density = c(1, 1, 1))
  

```

Use these initial parameters to generate simulated data parameters. Adding noise to the germination, survival, and lambda (max fecundity) parameters and making sure all parameters are non-negative. Also adding noise to the growth rates generated above (note that these are based on the average parameter values and NOT specifically connected to the noise in the individual parameters we just generated). Put in 0s instead of NAs because it was having issues, so we change this after adding noise to the data. Alphas are staying constant across the dataset for now. 
```{r sim-data}

add.noise <- function(x) {
  n <-  length(x)
  noise <- rnorm(n, mean = x, sd = x/2)
  return(noise)
}

truncate.positive <- function(x){
  ifelse(x < 0, 0, x)
}

# germination does not differ by environmental condition or competition under these assumptions
plot.germ <- function(species, bg, env){
  
  if(species == 'av'){
    germ <- g.av  %>% add.noise() %>% truncate.positive()
  
  } else if(species == 'ss') {
    germ <- g.ss  %>% add.noise() %>% truncate.positive()
    
  } else { # sp
    germ <- NA
  } 
  return(germ)
}

# survival differs by environmental condition (ss and sp) and competition (ss)
plot.surv <- function(species, bg, env){
  
  pop.intra <- focal.density %>% with(density[focal == species])
  dens <- filter(background.density, background == bg)
  dens[,species] <- dens[,species] + pop.intra
  
  if(species == 'ss'){
    comp.ss <- 1 + a.ss.av * g.av * dens$av + a.ss.ss * g.ss * dens$ss + a.ss.sp * dens$sp
     if (env == 'wet'){
      surv <- (s.ss.wet/comp.ss) 
    }
    else{
      surv <- (s.ss.dry/comp.ss) 
    }
    surv <- surv %>% add.noise() %>% truncate.positive()
  
  } else if(species == 'sp') {
     if (env == 'wet'){
      surv <- s.sp.wet
    }
    else{
      surv <- s.sp.dry
    }
    surv <- surv %>% add.noise() %>% truncate.positive()
    
  } else { # av
    surv <- NA
  } 
  return(surv)
}

# fecundity differs by environmental condition (av and sp) and competition (av and sp)
plot.fec <- function(species, bg, env){
  
  pop.intra <- focal.density %>% with(density[focal == species])
  dens <- filter(background.density, background == bg)
  dens[,species] <- dens[,species] + pop.intra
  
  if(species == 'av'){
    comp.av <- 1 + a.av.av * g.av * dens$av + a.av.ss * g.ss * dens$ss + a.av.sp * dens$sp
     if (env == 'wet'){
      fec <- (lambda.av.wet)/ comp.av 
    }
    else{
      fec <- (lambda.av.dry)/ comp.av 
    }
    fec <- fec %>% add.noise() %>% truncate.positive()
  
  } else if(species == 'sp') {
    comp.sp <- 1 + a.sp.av * g.av * dens$av + a.sp.ss * g.ss * dens$ss + a.sp.sp * dens$sp
     if (env == 'wet'){
      fec <- (lambda.sp.wet)/ comp.sp 
    }
    else{
      fec <- (lambda.sp.dry)/ comp.sp 
    }
    fec <- fec %>% add.noise() %>% truncate.positive()
    
  } else { # ss
    fec <- NA
  } 
  return(fec)
}

data.sim.2 <- data.sim

data.sim.2$germ<- pmap(list(data.sim$focal, data.sim$background, data.sim$water), plot.germ) %>% unlist()
data.sim.2$surv<- pmap(list(data.sim$focal, data.sim$background, data.sim$water), plot.surv) %>% unlist()
data.sim.2$fec<- pmap(list(data.sim$focal, data.sim$background, data.sim$water), plot.fec) %>% unlist()


```



## Estimate parameters from simulated data
Using a Bayesian model and rstan to estimate alpha parameters under different environmental conditions. What do we put for the number of competitors? The full number in the plot, or some relevant number of neighbors? 

We would want to check the data in a simple way (ex anova) to see which life stages were sensitive to 1. env 2. competition before formalizing that in the stan model -- here we're skipping that step and going with the way I set it up in the simulations. 

```{r model-parameters}
remove.na <- function(x){
  if_else(condition = is.na(x), 
          true = 0, 
          false = as.numeric(x))
}


data.sim.3 <- data.sim.2 %>% left_join(background.density) %>%
  mutate(ss = if_else(focal == 'ss', ss + 1, ss)) %>%
  mutate(sp = if_else(focal == 'sp', sp + 1, sp)) %>%
  mutate(av = if_else(focal == 'av', av + 1, av)) %>%
  mutate(focal_ss = if_else(focal == 'ss', 1, 0)) %>%
  mutate(focal_sp = if_else(focal == 'sp', 1, 0)) %>%
  mutate(focal_av = if_else(focal == 'av', 1, 0)) %>% 
  mutate(across(germ:fec, remove.na)) %>% # want to figure out a way to do this that we don't need this
  mutate(wet = if_else(water == 'wet', 1, 0))

stan.data <- data.sim.3 %>% 
  select(focal_ss, focal_sp, focal_av, wet, av, ss, sp, germ, surv, fec) %>% 
  compose_data()

```



## Unused

Growth rates are based on the average parameter values and NOT specifically connected to the noise in the individual parameters we just generated. I think this actually points out that growth rates won't be associated with a single plot, at least not for the ss and sp values, since they're a combination of values from each of those life stages and not just a pairwise combination. So we may actually only generate average growth rates for each life stage x environmental condition, rather than a unique one for each plot/replicate? But that's not right either because growth rates do differ by background competition. And we can use the relevant SP value from that same background condition/treatment/replicate for each one, or make a distribution of values across the replicates, rather than just taking the average. 

SO it would be great to stop and think about how we're actually going to calculate growth rates from the data we collect

```{r unused}
pop.growth <- function(species, bg, env){
  
  pop.intra <- focal.density %>% with(density[focal == species])
  dens <- filter(background.density, background == bg)
  dens[,species] <- dens[,species] + pop.intra
  
  if(species == 'av'){
    comp.av <- 1 + a.av.av * g.av * dens$av + a.av.ss * g.ss * dens$ss + a.av.sp * dens$sp
    if (env == 'wet'){
      growth <- lambda.av.wet / comp.av
    }
    else{
      growth <- lambda.av.dry / comp.av
    }
  
  } else if(species == 'ss') {
    comp.ss <- 1 + a.ss.av * g.av * dens$av + a.ss.ss * g.ss * dens$ss + a.ss.sp * dens$sp
    if (env == 'wet'){
      growth <- lambda.sp.wet / comp.ss
    }
    else{
      growth <- lambda.sp.dry / comp.ss
    }
    
  } else { # sp
    comp.sp <- 1 + a.sp.av * g.av * dens$av + a.sp.ss * g.ss * dens$ss + a.sp.sp * dens$sp
    if (env == 'wet'){
      pop.end <- (dens$ss * g.ss * s.ss.wet / comp.sp) + (dens$sp * s.sp.wet)
      growth <- pop.end / dens$sp
    }
    else{
      pop.end <- (dens$ss * g.ss * s.ss.dry / comp.sp) + (dens$sp * s.sp.dry)
      growth <- pop.end / dens$sp
    }
  } 
  return(growth)
}

data.sim$growth <- pmap(list(data.sim$focal, data.sim$background, data.sim$water), pop.growth) %>% unlist()


remove.na <- function(x){
  if_else(condition = is.na(x), 
          true = 0, 
          false = as.numeric(x))
}


data.sim.2 <- data.sim %>% 
  left_join(parameters, c('water', 'focal' = 'species')) %>%
  mutate(across(germ:lambda, remove.na)) %>%
  mutate(across(germ:lambda, add.noise)) %>%
  mutate(across(germ:lambda, truncate.positive))

 data.sim.2[data.sim.2$focal == 'sp','germ'] <- NA
 data.sim.2[data.sim.2$focal == 'av','surv'] <- NA
 data.sim.2[data.sim.2$focal == 'ss','lambda'] <- NA
```

