---
title: "partitioning_general"
output: html_document
---
Goal is to look at the strength of the storage effect in seedbanking annual (AS) and nonseedbanking annual (AN) species interactions. We want to measure how the storage effect [and the relative strength of storage effect:relative nonlinearity] changes with 1. magnitude of envrionmental variation 2. autocorrelation of environmental variation *This is the testing version to figure out what's going on with the tri-modal pattern in the resident mean log growth rates*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(xtable)
library(here)

## Personal ggplot theme
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
    #  strip.text=element_text(size=12),
    #  axis.text=element_text(size=12),
      complete = TRUE
    )
}
```

## Parameters and set up

This step defines which coexistence pair we're looking at, seedbanking annual with non-seedbanking annual, or perennial with non-seedbanking annual. Then sets parameters we're going to use for alpha and lambda under different environmental conditions. 

For the seedbanking annual with non-seedbanking annual interaction, uses lambda and alpha values for avena (as our AN) and erodium (as our AS) from Hallett et al. 2019 Ecology Letters paper, using the "wet" treatment as our wet environment and the "dry" treatment for our dry environment. Also going to use the seedbank parameters for as listed in that ms, which references Rice 1985, and simplifying the an to being non-seedbanking (so shifting its survival in the seed bank proportion from 0.4 to 0) *Note: planning to simplify this to be more about theoretical expectations*

For the perennial with non-seedbanking annual interaction, uses the theoretical parameters. 
```{r parameters}

# seedbanking annual - nonseedbanking annual parameters
pars.as.an <- read.csv(here('model-parameters/model-parameters_as_lambda-same.csv')) %>%
  mutate(stem.surv = NULL) 

# perennial - nonseedbanking annual parameters
pars.p.an <- read.csv(here('model-parameters/model-parameters_p_lambda-same.csv')) %>%
  mutate(seed.surv = NULL)

```

Functions for one step of species interaction calculations. Trying to make it general to whether it's both species present or just one--if it's one, that population is just at 0. Takes in the full parameter df for all the species (just for one environmental condition though) and a list of species population sizes. If it's just one species running to equilibrium then it's N0 will be 0. 
```{r species-interactions}
pop_interactions <- function(N0, param){
  
  # seedbanking annual - nonseedbanking annual
  if(pair == 'as.an'){
    pars.as <- param %>% filter(species == 'as')
    pars.an <- param %>% filter(species == 'an')
    
    n.as <- N0$as * (1 - pars.as$seed.germ) * pars.as$seed.surv + 
      N0$as * pars.as$seed.germ *pars.as$lambda /
      (1 + pars.as$a.an * pars.an$seed.germ * N0$an +
         pars.as$a.as * pars.as$seed.germ * N0$as)
    
    n.an <- N0$an * pars.an$seed.germ *pars.an$lambda /
      (1 + pars.an$a.an * pars.an$seed.germ * N0$an +
         pars.an$a.as * pars.as$seed.germ * N0$as)
    
    n.all <- tibble(an = n.an,
                    as = n.as)
  }
  
  # perennial - nonseedbanking annual
  if(pair == 'p.an'){
    pars.s <- param %>% filter(species == 's')
    pars.p <- param %>% filter(species == 'p')
    pars.an <- param %>% filter(species == 'an')
    
    
  n.s <- N0$p*pars.p$lambda /
    (1 + pars.p$a.an * pars.an$seed.germ * N0$an +
       pars.p$a.s * pars.s$seed.germ * N0$s +
       pars.p$a.p * N0$p)
  
  n.p <- N0$p * pars.p$stem.surv +
    (N0$s * pars.s$seed.germ * pars.s$stem.surv) /
    (1 + pars.s$a.an * pars.an$seed.germ * N0$an +
       pars.s$a.s * pars.s$seed.germ * N0$s +
       pars.s$a.p * N0$p)
  
  n.an <- N0$an * pars.an$seed.germ *pars.an$lambda /
    (1+ pars.an$a.an * pars.an$seed.germ * N0$an +
       pars.an$a.s * pars.s$seed.germ * N0$s +
       pars.an$a.p * N0$p)
    
    n.all <- tibble(an = n.an,
                    s = n.s,
                    p = n.p)
  }
  return(n.all)
}

```

## Running models

Partition to run the invasion, starting with one resident species (sp.resident) running on its own for init.time.steps starting at a population size of N0.r, and then invading sp.invader in from a total population size of 1 at each of the remaining time steps up to total.time.steps. Takes in a vector env.cond of environmental conditions. Currently only supports 2 different environmental states. 

Coded so that sp.invader and sp.resident parameters can be either a single value (ex: 'a') for a species with a single life stage or multiple values (ex: c('s','p')) for a species with multiple life stages. 

Has the options to vary lambda parameters with the environment (verus using the average), vary alpha parameters with the environment (versus using the average), and allowing lambda and alpha to covary (vs being sampled independently). For partitions where lambda is set to not vary, also sets germination and survival terms to not vary.
```{r running-invasion}
run_invasion <- function(sp.invader, sp.resident, N0.r, parameters, env.cond,
                         vary.lambda = TRUE, vary.alpha = TRUE, covary = TRUE,
                         init.time.steps = 50, total.time.steps = 100) {
  
  
  # populations with resident species not impacted by the invader
  # NOTE would like to change this to be more general 
  
  if(pair == 'as.an'){
    n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                    as = as.numeric(rep(NA, total.time.steps)))
  } 
  else if(pair == 'p.an'){
    n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                    s = as.numeric(rep(NA, total.time.steps)),
                    p = as.numeric(rep(NA, total.time.steps)))
  }
  
  # setting initial populations
  n.all[1,sp.resident] <- N0.r
  n.all[1,sp.invader] <- 0
  
  
  # name of column(s) for resident species pop when impacted by invading pop
  sp.res.impact <- paste(sp.resident, 'impact', sep = '.')
  
  # parameters for the average env condition
  p.ave <- parameters %>% filter(treatment == 'average')
  
  # columns for the non-varying parameters
  alpha.columns <- c(paste('a', sp.invader, sep = '.'), 
                     paste('a', sp.resident, sep = '.'))
  lambda.columns <- parameters %>% 
    select(!c(all_of(alpha.columns),'species','treatment')) %>% 
    names()
  
  # environmental condition vectors for the no covariance option
  env.lambda <- sample(env.cond, size = length(env.cond), replace = FALSE)
  env.alpha <- sample(env.cond, size = length(env.cond), replace = FALSE)
  
  for(t in 1:init.time.steps){
    # parameters for the varying env condition
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
    # constant lambda
    if(vary.lambda == FALSE){
      p.env[,lambda.columns] <- p.ave[,lambda.columns]
    }
    
    # constant alpha
    if(vary.alpha == FALSE){
      p.env[,alpha.columns] <- p.ave[,alpha.columns]
    }
    
    # parameters for no covariance between alpha and lambda
    if(covary == FALSE){
      lambda.env  <- parameters %>% filter(treatment == env.lambda[t])
      alpha.env <- parameters %>% filter(treatment == env.alpha[t])
      p.env[,lambda.columns] <- lambda.env[,lambda.columns]
      p.env[,alpha.columns] <- alpha.env[,alpha.columns]
    }
    # running one time step with the appropriate environmental parameters
    n.all[t+1,] <- pop_interactions(N0 = n.all[t,], param = p.env)
    
  }
  
  
  # figuring out rough invader population distribution if the invader is perennial
  # using average parameters and resident population size during the final warm-up step
  # makes it faster for running in for loop
  if(length(sp.invader) > 1) {

    n.inv.start <- 1/length(sp.invader)
    n.old <- n.all[init.time.steps,]  ## NOTE THIS IS FROM A SINGLE TIME POINT NOT THE AVERAGE
    n.old[sp.invader] <- n.inv.start
    n.old.inv <- n.inv.start[1]
    change <- 1

    while(change > 0.1){
      n.update <- pop_interactions(N0 = n.old, param = p.ave)

      n.per.total <- sum(n.update[sp.invader])
      n.update[sp.invader] <- n.update[sp.invader]/n.per.total
      n.update[sp.resident] <- n.old[sp.resident]
      n.old <- n.update
      n.new.inv <- n.update[sp.invader]
      change <- abs(n.old.inv - n.new.inv[1])
      n.old.inv <- n.new.inv[1]
    }
  }
  
  # running the resident and invader for the remaining time steps
  for (t in (init.time.steps+1):(total.time.steps-1)){
    # parameters for the varying env condition
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
   # adjusting non-varying or non-covarying parameters
    
    # constant lambda
    if(vary.lambda == FALSE){
      p.env[,lambda.columns] <- p.ave[,lambda.columns]
    }
    
    # constant alpha
    if(vary.alpha == FALSE){
      p.env[,alpha.columns] <- p.ave[,alpha.columns]
    }
    
    # parameters for no covariance between alpha and lambda
    if(covary == FALSE){
      lambda.env  <- parameters %>% filter(treatment == env.lambda[t])
      alpha.env <- parameters %>% filter(treatment == env.alpha[t])
      p.env[,lambda.columns] <- lambda.env[,lambda.columns]
      p.env[,alpha.columns] <- alpha.env[,alpha.columns]
    }
    
    
    # running invader and resident together
    N0.together <- n.all[t,]
    
    # figuring out starting invader population distribution if the invader is perennial
    # starting from the n.old found for average conditions above
     if(length(sp.invader) > 1){
      change <- 1
      n.old[sp.resident] <- N0.together[sp.resident]
      
      while(change > 0.01){
        n.update <- pop_interactions(N0 = n.old, param = p.env)
        n.per.total <- sum(n.update[sp.invader])
        n.update[sp.invader] <- n.update[sp.invader]/n.per.total
        n.update[sp.resident] <- N0.together[sp.resident] # MAYBE THIS IS THE ISSUE--SHOULDN'T BE FROM AVERAGE BUT FROM THAT TIME POINT
        n.new.inv <- n.update[sp.invader]
        change <- abs(n.old.inv - n.new.inv[1])
        n.old <- n.update
        n.old.inv <- n.new.inv[1]
      }
      N0.together[sp.invader] <- n.update[sp.invader] # using the equilibrium stage distribution from above
    } else {
      N0.together[sp.invader] <- 1 # if the invader has only one life stage
    }
    
    # running one time step with the appropriate environmental parameters
    n.res.impact <- pop_interactions(N0 = N0.together, param = p.env)
    
    # saving this output as the invader population and the impacted resident population
    n.all[t+1,sp.invader] <- n.res.impact[sp.invader]
    n.all[t+1,sp.res.impact] <- n.res.impact[sp.resident]
    
    # a version of this with the resident unaffected by the invader
    N0.res.only <- n.all[t,] %>% select(all_of(sp.invader), all_of(sp.resident))
    N0.res.only[sp.invader] <- 0
    n.res.only <- pop_interactions(N0 = N0.res.only, param = p.env)
    n.all[t+1,sp.resident] <- n.res.only[sp.resident]
    
    # growth rate of the resident
    sp.res.gr <- paste('gr', paste(sp.resident, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.res.gr] <- sum(n.all[t+1,sp.res.impact])/sum(n.all[t,sp.resident])
    
    # growth rate of the invader
    sp.inv.gr <- paste('gr', paste(sp.invader, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.inv.gr] <- sum(n.all[t+1,sp.invader])
    
  }
  
  n.all$time <- 1:total.time.steps
  return(n.all)
  
}

```

Plotting function so we can visualize what's going on with the different types of variation
```{r plotting-function}
plot_invasion <- function(df){ # plots actual population sizes
  df.long <- df %>% 
    pivot_longer(
      cols = !c(time, env, inv, res, run, starts_with('gr')),
                 names_to = 'species', values_to = 'n',
                 values_drop_na = TRUE) %>% 
    mutate(n.log = log(n))

  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    theme_cw() +
    xlab('Time') + 
    ylab('Population size (log)') +
    ylim(c(-5, 8))
  
}

plot_gr <- function(df) { # plots population growth rates
  df.long <- df %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', values_to = 'n',
                 values_drop_na = TRUE) %>% 
    mutate(n.log = log(n))
  
  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    geom_point() +
    theme_cw() +
    xlab('Time') + 
    ylab('Growth Rate (log)')
}
```

# Running and plotting invasions 
*Don't re-run this, just skip to the next step to load it*
Full variation with multiple runs for troubleshooting (no partitioning). This saves at the end and loads at the next step so we don't have to rerun it for time. 
```{r full-multiple-runs}

time.warm.up <- 50
time.full <- 400

case <- 'p'

pair <- paste(case, 'an', sep = '.')

# select the relevant parameter set
if(pair == 'as.an') {
  sp.1 <- 'an'
  sp.2 <- 'as'
  sp.order <- c('an','as') 
  pars <- pars.as.an 
}

if(pair == 'p.an'){
  sp.1 <- 'an'
  sp.2 <- c('s','p')
  sp.order <- c('an','s','p')
  pars <- pars.p.an
}

  sp.1.name <- paste(sp.1, collapse = '.')
  sp.2.name <- paste(sp.2, collapse = '.')
  
invade.runs <- tibble()

env.ratio <- 0.5

pars <- pars %>% mutate(weight = ifelse(treatment == 'wet', env.ratio, 1-env.ratio))

pars.ave <- pars %>% 
    group_by(species) %>%
    summarise_if(is.numeric,
                 ~ weighted.mean(., weight)) %>%
  mutate(weight = NA, treatment = 'average') %>% 
  arrange(factor(species, levels = sp.order))  # need to keep them in the right order for later

for(i in 1:20){
  # two equally probable environmental conditions
env.draw <- rbinom(time.full, 1, env.ratio) 
env.condition <- if_else(env.draw > 0, 'wet','dry')

pars <- pars %>% full_join(pars.ave)

invade.1 <-  run_invasion(sp.invader = sp.1, sp.resident = sp.2, N0.r = 100, 
                         parameters = pars, env.cond = env.condition,
                         vary.lambda = TRUE, vary.alpha = TRUE, 
                         init.time.steps = time.warm.up, total.time.steps = time.full) %>%
  add_column(inv = sp.1.name, res = sp.2.name, run = i, env = env.condition) %>%
  filter(time > time.warm.up + 1)

invade.2 <-  run_invasion(sp.invader = sp.2, sp.resident = sp.1,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = TRUE, vary.alpha = TRUE) %>%
  add_column(inv = sp.2.name, res = sp.1.name, run = i, env = env.condition) %>%
  filter(time > time.warm.up + 1)


invade.both <- bind_rows(invade.1, invade.2)

invade.runs <- bind_rows(invade.runs, invade.both)
}

# save(invade.runs, file = "testing_trimodal.Rdata")

```


Loading in the file from above so we can run tests on it. One note: the longer I run the runs for (number of time steps) the closer the two "wings" get to 0. So for 200 time steps they were around plus/minus 0.015, and with 400 time steps they're around plus/minus 0.005. On a separate test I ran each one for 1000 time steps and it's around plus/minus 0.002, but still clearly tri-modal. 

```{r testing-trimodal}

load(file = "testing_trimodal.Rdata")
#load(file = "testing_trimodal_as.Rdata")

sp.res.gr <- invade.runs %>% group_by(inv, res, run) %>%
  dplyr::summarise(
    gr.an.log.mean = mean(log(gr.an)),
    gr.an.mean = mean(gr.an)
  ) %>% filter(res == 'an')

hist(sp.res.gr$gr.an.mean) # not tri-modal if we just take the means
hist(sp.res.gr$gr.an.log.mean) # tri-modal if we take the means of the logs
which(sp.res.gr$gr.an.log.mean > 0.004)
which(sp.res.gr$gr.an.log.mean < -0.004)


# plotting options that have mean growth rates in the different groups
# numbers would need to be changed for the an-as comparison
# based on above results of which is in which group
test.1 <- filter(invade.runs, run == 1, res == 'an')
test.3 <- filter(invade.runs, run == 3, res == 'an')
test.11 <- filter(invade.runs, run == 11, res == 'an')

# plotting log growth rates -- nothing seems to be different between these different runs
plot_gr(test.1)
plot_gr(test.3)
plot_gr(test.11)

# looking at the histograms of the growth rates (not logged) for these different ones
# looks bi-modal when generalized but with finer breaks is actually split into
# 4 sections. This seems to make sense matching the patterns of growth we see in the graphs
hist(test.1$gr.an, ylim = c(0, 250), breaks = 30) 
hist(test.3$gr.an, ylim = c(0, 250), breaks = 30)
hist(test.11$gr.an, ylim = c(0, 250), breaks = 30)

# log growth rates within individual runs are generally tri-modal
# although with finer breaks again it may actually be 5 sections
hist(log(test.1$gr.an), ylim = c(0, 250), breaks = 30) 
hist(log(test.3$gr.an), ylim = c(0, 250), breaks = 30)
hist(log(test.11$gr.an), ylim = c(0, 250), breaks = 30)

mean(log(test.1$gr.an)) # high
mean(log(test.3$gr.an)) # near 0
mean(log(test.11$gr.an)) # low

# no clear difference between these 3 that would be causing the difference in means,
# other than there just being more or fewer of the different growth rate groups?

mean(log(test.1[test.1$time > 200,]$gr.an)) # high
mean(log(test.3[test.3$time > 200,]$gr.an)) # low
mean(log(test.11[test.11$time > 200,]$gr.an)) # near 0

# the means change if we subset the data, with test.3 and test.11 switching groups
# I still don't understand why it doesn't end up being uni-modal

```
What about when the perennial species is the resident?
```{r testing-trimodal-sp}
load(file = "testing_trimodal.Rdata")

sp.res.gr <- invade.runs %>% group_by(inv, res, run) %>%
  dplyr::summarise(
    gr.sp.log.mean = mean(log(gr.s.p)),
    gr.sp.mean = mean(gr.s.p),
  ) %>% filter(res == 's.p')

hist(sp.res.gr$gr.sp.mean) # not tri-modal if we just take the means
hist(sp.res.gr$gr.sp.log.mean) # trimodal

```


What about when the seedbanking annual is the resident?
```{r testing-trimodal-sp}
load(file = "testing_trimodal_as.Rdata")

sp.res.gr <- invade.runs %>% group_by(inv, res, run) %>%
  dplyr::summarise(
    gr.as.log.mean = mean(log(gr.as)),
    gr.as.mean = mean(gr.as),
  ) %>% filter(res == 'as')

hist(sp.res.gr$gr.as.mean) # not tri-modal if we just take the means
hist(sp.res.gr$gr.as.log.mean) # NOT tri-modal

# so what's different here? 
test.1 <- filter(invade.runs, run == 1, res == 'as')
plot_gr(test.1)
hist(log(test.1$gr.as), ylim = c(0, 250), breaks = 0) 
# individual time step growth rates are not multi-modal either

```

# NOT USED -- ISSUES ARE IN FULL RUNS, BEFORE THE PARTITION STEPS

# Calculating partitions 
Function to calculate partitions for a given invader and resident
```{r partition-function}
partition_epsilons <- function(sp.inv, sp.res, pars, 
                               env.condition, time.warm.up, time.full){
  res.name <- paste(sp.res, collapse = '.')
  inv.name <- paste(sp.inv, collapse = '.')
  
  part <- tibble(sp.invader = inv.name, sp.resident = res.name)
  
  # no variation
  invade.none <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, 
                               total.time.steps = time.full,
                               vary.lambda = FALSE, vary.alpha = FALSE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  part$inv.e0 <- invade.none %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  part$res.e0 <- invade.none %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  # lambda variation only
  invade.l <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                            N0.r = 100, parameters = pars, env.cond = env.condition,
                            init.time.steps = time.warm.up, total.time.steps = time.full,
                            vary.lambda = TRUE, vary.alpha = FALSE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  l.inv.mean <- invade.l %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  l.res.mean <- invade.l %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.el <- l.inv.mean - part$inv.e0 
  part$res.el <- l.res.mean- part$res.e0
  
  # alpha variation only
  invade.a <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                            N0.r = 100, parameters = pars, env.cond = env.condition,
                            init.time.steps = time.warm.up, total.time.steps = time.full,
                            vary.lambda = FALSE, vary.alpha = TRUE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  a.inv.mean <- invade.a %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  a.res.mean <- invade.a %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.ea <- a.inv.mean - part$inv.e0 
  part$res.ea <- a.res.mean- part$res.e0
  
  # full variation
  invade.full <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = TRUE, vary.alpha = TRUE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  part$inv.full <- invade.full %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  part$res.full <- invade.full %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.eint <- part$inv.full - (part$inv.e0 + part$inv.el + part$inv.ea)
  part$res.eint <- part$res.full - (part$res.e0 + part$res.el + part$res.ea)
  
  # no correlation
   invade.nocov <- run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.r = 100, parameters = pars, env.cond = env.condition,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = TRUE, vary.alpha = TRUE,
                               covary = FALSE) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
   
   nocov.inv <- invade.nocov %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
   nocov.res <- invade.nocov %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
   
   part$inv.nocov <- nocov.inv - (part$inv.e0 + part$inv.el + part$inv.ea)
   part$res.nocov <- nocov.res - (part$res.e0 + part$res.el + part$res.ea)
   
   # storage effect
   part$inv.storage <- part$inv.eint - part$inv.nocov
   part$res.storage <- part$res.eint - part$res.nocov
  
  return(part)
}


```

Running that partition function *This would get split into a separate script to run on Teton computing cluster*
```{r partition-run}

# CHOOSE: This determines which coexistence pair we're looking at
#case <- 'as'
 case <- 'p'

pair <- paste(case, 'an', sep = '.')

varying <- 'lambda'
#varying <- 'alpha'

favorable <- 'same'
# favorable <- 'same-consistent'
# favorable <- 'opposite-consistent'
# favorable <- 'same-flipped'
# favorable <- 'opposite-flipped'

# file names

file.core <- paste0(case, '_', varying, '-', favorable)
main.wd <- '/project/coexistence/cwerner5/'  # needs to be updated

file.in <- paste0(main.wd, 'model-parameters/model-parameters_', file.core, '.csv')

file.out <- paste0(main.wd, 'model-data/mct-partitions_', file.core, '.RData')

# SET PARAMETER VALUES 
pars <- read.csv(file.in) %>% # THIS NEEDS TO BE GENERALIZED
  mutate(stem.surv = NULL) 


# select the relevant parameter set
if(pair == 'as.an') {
  sp.1 <- 'an'
  sp.2 <- 'as'
  sp.order <- c('an','as') 
}
if(pair == 'p.an'){
  sp.1 <- 'an'
  sp.2 <- c('s','p')
  sp.order <- c('an','s','p')
}

env.ratio <- 0.5 # ratio of wet years to dry years

# adding rows for the weighted average of the parameters
pars <- pars %>% mutate(weight = ifelse(treatment == 'wet', env.ratio, 1-env.ratio))

pars.ave <- pars %>% 
    group_by(species) %>%
    summarise_if(is.numeric,
                 ~ weighted.mean(., weight)) %>%
  mutate(weight = NA, treatment = 'average') %>% 
  arrange(factor(species, levels = sp.order)) # need to keep them in the right order for later

pars <- pars %>% full_join(pars.ave)

time.warm.up <- 200
time.full <- 300

# Next steps: Repeat this multiple times! Start trying out different scenarios!
# Also probably get it set up to work on Teton to save time

# Environmental conditions
env.draw <- rbinom(time.full, 1, env.ratio) # sequence of the environment (NOTE: would want this to change if doing multiple runs)
env.condition <- if_else(env.draw > 0, 'wet','dry')

partitions <- partition_epsilons(sp.inv = sp.2, sp.res = sp.1, pars, 
                                 env.condition, time.warm.up, time.full) %>%
  rbind(partition_epsilons(sp.inv = sp.1, sp.res = sp.2, pars, 
                           env.condition, time.warm.up, time.full)) %>%
  #pivot_longer(cols = inv.e0:res.eint) %>%
  pivot_longer(cols = inv.e0:res.storage) %>%
  separate(name, into = c('player', 'partition')) %>%
  pivot_wider(names_from = player, values_from = value) %>%
  mutate(delta = inv - res,
         partition = factor(partition, 
                            levels = c('full','e0','el','ea','eint',
                                       'storage','nocov')))

# Titles for invasion interactions
if(pair == 'as.an'){
  species_names <- c(
    an = 'AN invades AS',
    as = 'AS invades AN'
  )
  file.out <- 'partitions-AS-AN.pdf'
}
if(pair == 'p.an'){
  species_names <- c(
    an = 'AN invades P',
    s.p = 'P invades AN'
  )
  file.out <- 'partitions-P-AN.pdf'
}


cb_palette <- c("black", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")
colors2 <- c("#749dae", rep('darkslategrey', times = 4), rep("#f3c483", times = 2))

ggplot(partitions, aes(x = partition, y = delta, fill = partition)) + 
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_cw() +
  ylab('Partitioning of growth rate when rare') +
  scale_x_discrete(labels = c('full' = 'Full',  'e0' = expression(Delta^0),
                              'el' = expression(Delta^lambda), 'ea' = expression(Delta^alpha),
                              'eint' = expression(Delta^{alpha*lambda}),
                              'storage' = expression(Delta^(alpha*lambda)), 
                              'nocov' = expression(Delta^{alpha!=lambda})),
                   name = 'Partition') +
  scale_fill_manual(values = cb_palette) +
  guides(fill = 'none')

 # ggsave(file.out, width = 8, height = 5, units = 'in')

```

Not used: Running over an environmental sequence instead of a particular env.ratio. This was used for testing before running on the computer cluster
```{r env-seq}
# This determines which coexistence pair we're looking at
pair <- 'as.an'
#pair <- 'p.an'

# select the relevant parameter set
if(pair == 'as.an') {
  pars <- pars.as.an
  sp.1 <- 'an'
  sp.2 <- 'as'
  sp.order <- c('an','as') 
  
}
if(pair == 'p.an'){
  pars <- pars.p.an
  sp.1 <- 'an'
  sp.2 <- c('s','p')
  sp.order <- c('an','s','p')
}


# number of time steps for each run
time.warm.up <- 50
time.full <- 100

# initialise dataframe
  partitions.env <- tibble(sp.invader = character(), sp.resident = character(),
                           partition = character(),
                           inv = numeric(), res = numeric(), delta = numeric(), 
                           env.ratio = numeric())
  env.seq <- seq(0, 1, by = 0.25) # ratio of wet years to dry years
  
  for(env.ratio in env.seq){
    # adding rows for the weighted average of the parameters
    pars.2 <- pars %>% mutate(weight = ifelse(treatment == 'wet', env.ratio, 1 - env.ratio))
    
    pars.ave <- pars.2 %>% 
      group_by(species) %>%
      summarise_if(is.numeric,
                   ~ weighted.mean(., weight)) %>%
      mutate(weight = NA, treatment = 'average') %>% 
      arrange(factor(species, levels = sp.order)) # need to keep them in the right order for later
    
    pars.2 <- pars.2 %>% full_join(pars.ave)
    
    
    env.draw <- rbinom(time.full, 1, env.ratio) # sequence of the environment (NOTE: would want this to change if doing multiple runs)
    env.condition <- if_else(env.draw > 0, 'wet','dry')
    
    partitions <- partition_epsilons(sp.inv = sp.2, sp.res = sp.1, pars.2, 
                                     env.condition, time.warm.up, time.full) %>%
      rbind(partition_epsilons(sp.inv = sp.1, sp.res = sp.2, pars.2, 
                               env.condition, time.warm.up, time.full)) %>%
      #pivot_longer(cols = inv.e0:res.eint) %>%
      pivot_longer(cols = inv.e0:res.storage) %>%
      separate(name, into = c('player', 'partition')) %>%
      pivot_wider(names_from = player, values_from = value) %>%
      mutate(delta = inv - res,
             partition = factor(partition, 
                                levels = c('full','e0','el','ea','eint', 'storage','nocov')))
    partitions$env.ratio <- env.ratio
    
    partitions.env <- full_join(partitions.env, partitions, 
                                by = c('sp.invader','sp.resident','partition',
                                       'inv','res', 'delta','env.ratio'))
  }
  
  partitions.env <- partitions.env  %>%
    mutate(partition = factor(partition, 
                            levels = c('full','e0','el','ea','eint',
                                       'storage','nocov')))
  
  partition.legend <- c(
  'full' = 'Full',  
  'e0' = expression(Delta^0), 
  'el' = expression(Delta^lambda), 
  'ea' = expression(Delta^alpha),
  'eint' = expression(Delta^{alpha*lambda}),
  'storage' = expression(Delta^(alpha*lambda)), 
  'nocov' = expression(Delta^{alpha!=lambda})
)
  
  ggplot(partitions.env, aes(x = env.ratio, y = delta, color = partition)) +
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_line() +
  theme_cw() +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
  xlab ('Proportion wet years') +
  ylab('Delta')

```


# Loading in rdata file from partitions run on the computer cluster
```{r partition-seq}

# pair <- 'as.an'
# pair <- 'p.an'

# case <- 'as'
 case <- 'p'

# do only lambda/germination/survival terms vary by environment or do the alpha terms vary too?
 varying <- 'lambda'
# varying <- 'alpha'

# parameter settings for favorable vs unfavorable environments
# for lambda varying the options are 'same' or 'opposite'
# with alpha also varying the options are 'same-consistent', 'same-flipped',
# 'opposite-consistent', and 'opposite-flipped' where same and opposite determine whether
# the two species have similar or opposing responses to the environmental condition
# and consistent or flipped determine whether the lambda and alpha terms have similar
# or opposing responses to the environment

 favorable <- 'same'
# favorable <- 'opposite'
# favorable <- 'same-consistent' 
# favorable <- 'same-flipped' 
# favorable <- 'opposite-consistent' 
# favorable <- 'opposite-flipped'


file.core <- paste0('mct-partitions_', case, '_', varying, '-', favorable)

load(here(paste0('model-data/', file.core, '.RData')))
plot.out.env <- here(paste0('figures/', file.core, '_env-ratio.pdf')) 
plot.out.50 <- here(paste0('figures/', file.core, '_50.pdf'))

# Load data and set titles for invasion interactions
if(case == 'as'){
  species_names <- c(
    an = 'AN invades AS',
    as = 'AS invades AN'
  )
}
if(case == 'p'){
  species_names <- c(
    an = 'AN invades P',
    s.p = 'P invades AN'
  )
}

# summary of the partition results
part.long <- simulations %>% map_dfr(bind_rows, .id = "run") %>%
  mutate(partition = factor(partition, 
                            levels = c('full','e0','el','ea','eint',
                                       'storage','nocov')))

part.sum <- part.long %>% 
  group_by(sp.invader, sp.resident, partition, env.ratio) %>%
  dplyr::summarise(
    inv.mean = mean(inv),
    res.mean = mean(res),
    delta.mean = mean(delta),
    delta.sd = sd(delta) # want to switch this to CI approach?
  )

# plot stuff
partition.legend <- c(
  'full' = 'Full',  
  'e0' = expression(Delta^0), 
  'el' = expression(Delta^lambda), 
  'ea' = expression(Delta^alpha),
  'eint' = expression(Delta^{alpha*lambda}),
  'storage' = expression(Delta^(alpha*lambda)), 
  'nocov' = expression(Delta^{alpha!=lambda})
)

cb_palette <- c("black", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")

# plotting across the ratio of environmental conditions
ggplot(part.sum, aes(x = env.ratio)) +
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_ribbon(alpha = 0.3, aes(ymin = delta.mean - delta.sd, 
                               ymax = delta.mean + delta.sd,
                               fill = partition)) +
  geom_line(aes(y = delta.mean, color = partition)) +
  theme_cw() +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
  xlab ('Ratio of env 1 to env 2') +
  ylab('Delta')

#  ggsave(plot.out.env, width = 8, height = 5, units = 'in')

# plotting for just the 50-50 condition
ggplot(filter(part.sum, env.ratio == 0.5), 
       aes(x = partition, y = delta.mean, 
           ymin = delta.mean - delta.sd, 
                               ymax = delta.mean + delta.sd,
           fill = partition)) + 
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_errorbar(position = position_dodge(), color = 'grey50', width = 0.2) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_cw() +
  ylab('Partitioning of growth rate when rare') +
  scale_x_discrete(labels = c('full' = 'Full',  'e0' = expression(Delta^0),
                              'el' = expression(Delta^lambda), 'ea' = expression(Delta^alpha),
                              'eint' = expression(Delta^{alpha*lambda}),
                              'storage' = expression(Delta^(alpha*lambda)), 
                              'nocov' = expression(Delta^{alpha!=lambda})),
                   name = 'Partition') +
  scale_fill_manual(values = cb_palette) +
  guides(fill = 'none')
#ggsave(plot.out.50, width = 8, height = 5, units = 'in')


# hist for the 50-50 condition
ggplot(filter(part.long, env.ratio == 0.5, partition == 'full', sp.invader == 'as'), 
       aes(inv)) +
  geom_histogram()

ggplot(filter(part.long, env.ratio == 0.5, partition == 'full', sp.invader == 'as'), 
       aes(res)) +
  geom_histogram()
  
  
### Plots to try to investigate the high variation in when the perennial is the invader
plot.out.sp.delta <- here(paste0('figures/', file.core, '_delta.pdf'))
plot.out.sp.inv <- here(paste0('figures/', file.core, '_inv.pdf'))
plot.out.sp.res <- here(paste0('figures/', file.core, '_res.pdf'))

## Plotting individual runs (delta)
ggplot(filter(part.long, run %in% unique(part.long$run)[1:20]), 
       aes(x = env.ratio, y = delta, color = run)) +
  facet_grid(cols = vars(sp.invader), rows = vars(partition)) + 
  geom_line(alpha = 0.7) +
  theme_cw() +
#  scale_color_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
#  scale_fill_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
  xlab ('Ratio of env 1 to env 2') +
  ylab('Delta') +
  theme(legend.position = 'none')
ggsave(plot.out.sp.delta, width = 6, height = 8, units = 'in')

## Plotting individual runs (resident)
ggplot(filter(part.long, run %in% unique(part.long$run)[1:20]), 
       aes(x = env.ratio, y = res, color = run)) +
  facet_grid(cols = vars(sp.invader), rows = vars(partition)) + 
  geom_line(alpha = 0.7) +
  theme_cw() +
#  scale_color_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
#  scale_fill_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
  xlab ('Ratio of env 1 to env 2') +
  ylab('Res') +
  theme(legend.position = 'none')
ggsave(plot.out.sp.res, width = 6, height = 8, units = 'in')

## Plotting individual runs (invader)
ggplot(filter(part.long, run %in% unique(part.long$run)[1:20]), 
       aes(x = env.ratio, y = inv, color = run)) +
  facet_grid(cols = vars(sp.invader), rows = vars(partition)) + 
  geom_line(alpha = 0.7) +
  theme_cw() +
#  scale_color_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
#  scale_fill_manual(values = cb_palette, labels = partition.legend, name = 'Partition') +
  xlab ('Ratio of env 1 to env 2') +
  ylab('Inv') +
  theme(legend.position = 'none')
ggsave(plot.out.sp.inv, width = 6, height = 8, units = 'in')
  
  
# table output to latex
# part.sum %>% 
#   filter(env.ratio %in% c(0.0, 0.25, 0.5, 0.75, 1.0)) %>%
#   arrange(sp.invader, env.ratio) %>%
#   xtable() %>%
#   print(., include.rownames = FALSE)

```

