---
title: "Partitioning perennial - nonseedbank annual"
output: html_document
---

Goal is to look at the strength of the storage effect in perennial (P) and nonseedbanking annual (AN) species interactions. We want to measure how the storage effect [and the relative strength of storage effect:relative nonlinearity] changes with 1. magnitude of envrionmental variation 2. autocorrelation of environmental variation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

## Personal ggplot theme
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
    #  strip.text=element_text(size=12),
    #  axis.text=element_text(size=12),
      complete = TRUE
    )
}
```

## Parameters and set up
Setting parameters we're going to use for alpha and lambda under different environmental conditions. 

```{r parameters}
# main parameters
pars <- read.csv('model_parameters_perennials.csv') %>%
  mutate(seed.surv = NULL)

# adding average values to the parameter df
an.means <- pars %>%
  filter(species == 'an') %>% 
  select(a.an, a.s, a.p, lambda, seed.germ) %>%
  colMeans()
s.means <- pars %>%
  filter(species == 's') %>% 
  select(a.an, a.s, a.p, seed.germ, stem.surv) %>%
  colMeans()
p.means <- pars %>%
  filter(species == 'p') %>%
  select(a.an, a.s, a.p, lambda, stem.surv) %>%
  colMeans()

pars <- pars %>%
  add_row(treatment = 'average', species = 'an', 
          a.an = an.means['a.an'], a.s = an.means['a.s'], a.p = an.means['a.p'],
          lambda = an.means['lambda'], seed.germ = an.means['seed.germ'], 
          stem.surv = NA) %>%
  add_row(treatment = 'average', species = 's', 
          a.an = s.means['a.an'], a.s = s.means['a.s'], a.p = s.means['a.p'],
          lambda = NA, seed.germ = s.means['seed.germ'], 
          stem.surv = s.means['stem.surv']) %>%
  add_row(treatment = 'average', species = 'p', 
          a.an = p.means['a.an'], a.s = p.means['a.s'], a.p = p.means['a.p'],
          lambda = p.means['lambda'], seed.germ = NA, 
          stem.surv = p.means['stem.surv'])

```

Functions for coexistence calculations. These are currently just for annual species and will need to change quite a bit--probably won't be able to just be general to resident/invaders but specific to which species is which?
```{r functions-coexistence}
# Determine equilibrium conditions for each species in isolation 
pop_equilibrium_annual <- function (N0, s, g, a.intra, lambda) {
  # to run for only a single timestep
  N <- s*(1-g)*N0 + N0*(lambda*g)/(1+a.intra*N0)
  return(N)
}

pop_equilibrium_perennial <- function (N0.s, N0.p, param) {
  # to run for only a single timestep
  pars.s <- param %>% filter(species == 's')
  pars.p <- param %>% filter(species == 'p')
  
  n.s <- N0.p*pars.p$lambda/(1 + pars.p$a.p*N0.p +
                               pars.p$a.s*pars.s$seed.germ*N0.s)
  
  n.p <- N0.p*pars.p$stem.surv +
    (N0.s*pars.s$seed.germ*pars.s$stem.surv)/
    (1 + pars.s$a)
  
}

# invader population growth rate one time step forward
pop_invade <- function (N0, resident, s, g, a.inter, lambda) {
  # to run for only a single timestep
  N <- s*(1-g)*N0 + N0*(lambda*g)/(1+a.inter*resident)
  return(N)
}

# resident population growth rate one time step forward
pop_resident <- function (N0, resident, s, g, a.intra, a.inter, lambda) {
  # to run for only a single timestep
  N <- s*(1-g)*resident + resident*(lambda*g)/(1+a.intra*resident+a.inter*N0)
  return(N)
}
```

## Running models
Function to run an invasion one way: resident only for init.time.steps, invader added in after that until total.time.steps. Returns a tibble with columns for species.
```{r running-invasion}
run_invasion <- function(sp.invader, sp.resident, N0.i, N0.r, parameters, vary.lambda = TRUE, vary.alpha = TRUE, init.time.steps = 50, total.time.steps = 100){
  
  # two equally probable environmental conditions
  env.draw <- rbinom(total.time.steps, 1, 0.5) 
  env.cond <- if_else(env.draw > 0, 'wet','dry')
  
  n.res <- rep(NA, total.time.steps) # resident population without the invader
  n.res.impact <- rep(NA, total.time.steps) # resident population with invader's impact
  n.res.gr <- rep(NA, total.time.steps) # growth rate of resident population (with invader's impact)
  n.inv <- rep(NA, total.time.steps) # invader population, init.time.steps will stay NA
  
  a.res <- paste('a',sp.resident, sep = '.') # column name of resident interactions
  a.inv <- paste('a',sp.invader, sep = '.') # column name of invader interactions
  
  # parameters for average environmental condition
  p.res.ave <- parameters %>% filter(species == sp.resident, treatment == 'average')
  p.inv.ave <- parameters %>% filter(species == sp.invader, treatment == 'average')
    
  # running just the resident alone for init.time.steps
  n.res[1] <- N0.r
  for (t in 1:init.time.steps){
    # parameters for varying environmental condition
    p.res <- parameters %>% filter(species == sp.resident, treatment == env.cond[t])
    
    # running one time step using the appropriate environmental parameters
    n.res[t+1] <- pop_equilibrium(N0 = n.res[t], 
                                  s = p.res$seed.surv, 
                                  g = p.res$seed.germ, 
                                  lambda = if_else(vary.lambda == TRUE,
                                                   p.res$lambda, 
                                                   p.res.ave$lambda), 
                                  a.intra = if_else(vary.alpha == TRUE,
                                                    p.res[,a.res],
                                                    p.res.ave[,a.res]))
    n.res.impact[t+1] <- n.res[t+1]
   #  n.res.gr[t+1] <- n.res[t+1]/n.res[t]
  }
  
  # running the resident and invader for the remaining time steps
  for (t in (init.time.steps+1):(total.time.steps-1)){
    p.res <- parameters %>% filter(species == sp.resident, treatment == env.cond[t])
    p.inv <- parameters %>% filter(species == sp.invader, treatment == env.cond[t])
    
    # invader population
    # N0 of the invader is set to the same value each time
    n.inv[t+1] <- pop_invade(N0 = N0.i, 
                             resident = n.res[t], #** Should be n.res[init.time.steps]
                             s = p.inv$seed.surv, 
                             g = p.inv$seed.germ, 
                             lambda = if_else(vary.lambda == TRUE,
                                              p.inv$lambda, 
                                              p.inv.ave$lambda), 
                             a.inter = if_else(vary.alpha == TRUE,
                                               p.inv[,a.res],
                                               p.inv.ave[,a.res]))
    
    # resident population as if invader weren't present
    n.res[t+1] <- pop_equilibrium(N0 = n.res[t], 
                                  s = p.res$seed.surv, 
                                  g = p.res$seed.germ, 
                                  lambda = if_else(vary.lambda == TRUE,
                                                   p.res$lambda, 
                                                   p.res.ave$lambda), 
                                  a.intra = if_else(vary.alpha == TRUE,
                                                    p.res[,a.res],
                                                    p.res.ave[,a.res]))
    
    # resident population with invader present
    n.res.impact[t+1] <- pop_resident(N0 = N0.i, 
                               resident = n.res[t], 
                               s = p.res$seed.surv, 
                               g = p.res$seed.germ, 
                               lambda = if_else(vary.lambda == TRUE,
                                                p.res$lambda, 
                                                p.res.ave$lambda), 
                               a.intra = if_else(vary.alpha == TRUE,
                                                 p.res[,a.res],
                                                 p.res.ave[,a.res]),
                               a.inter = if_else(vary.alpha == TRUE,
                                                 p.res[,a.inv],
                                                 p.res.ave[,a.inv]))
    n.res.gr[t+1] <- n.res.impact[t+1]/n.res[t]
  }

  return(tibble(time = 1:total.time.steps, env.cond, n.inv, n.res, n.res.impact, n.res.gr))
}

```

Additional function for partitioning the storage effect, using temporal randomization to remove correlations. Not entirely sure I'm doing this correctly but we're just going to go for it. Could include it in the function above but that would add a lot of nested if_else statements and I think this might be simpler--also easier to adjust based on feedback. 

*Question: should invader lambda and resident lambda stay correlated, ditto invader alpha and resident alpha? Or fully drop the covariance between all four of those? Currently trying the former*

*In the Ellner ForcedChemo_Func_Covar.R code, "exact sharping" uses all possible combinations of the two conditions (E = temp, C = resource) generated with the "expand.grid" function. And yes it looks like both species are run on the same temp and resource conditions for each step? r1 is for one species and r2 is for the other species, and V1/V2 is the lambda function. D is the dilution (outflow) rate--which is maybe equivalent to a death rate. But I'm not sure what K1 and K2 are--a function of the temperature, not of the resource--not the same as Q which is plotted in the paper. The ms says that K of each species is the "half-saturation constant" and is modeled as a constant but the code looks like it is a function of temperature? These r_j functions are (I think) not based on the density of either species? WHICH maybe makes sense because this isn't a partition involving competition*

*Okay, so first they do these simulations--one set with temp and R variation and one with no variation. minus1 is species1 is the invader species 2 is the resident, minus1$x1 is species1 in this setup. Resident effects on the invader are indirect, via the resident effects on the resource which are used to generate this time series. The only thing that changes between the full variation simulation and the no-variation simulation is parms2$a which goes from 6 to 0--this is the parameter that sets the amplitude of the temperature variations. Resource variation is driven by temperature variation effects on the resident species and its consequent resource uptake rather than set directly.*

*Something else to think about is the $\epsilon_1^{0}$ versus $\epsilon_1^{star}$ as the different base growth rates. I think we (and the Hallett 2019 paper) are using $\epsilon_1^{star}$, the completely fluctuation-independent growth rate. But Ellner has that as a part of $\epsilon_1^{0}$ which is what they then subtract out from the other terms --see the text below eqn 13*

*Also why do they use Esharp for calculating delta1.sigmaE (and ditto for Csharp/sigmaC) instead of just E and C the original time series. Did some quick calculations and it looks like it doesn't change anything--but does that mean that changing time series autocorrelation wouldn't change those components either?*

*Final question is whether the randomization approach we took is equivalent to the sharping or not, since those are based on manipulating the simulation time series after it's been run, not changing things up one step at a time*

```{r running-no-correlation}
run_invasion_no_correlation <- function(sp.invader, sp.resident, N0.i, N0.r, parameters, init.time.steps = 50, total.time.steps = 100){
  
  n.res <- rep(NA, total.time.steps) # resident population without the invader
  n.res.impact <- rep(NA, total.time.steps) # resident population with invader's impact
  n.res.gr <- rep(NA, total.time.steps) # growth rate of resident population (with invader's impact)
  n.inv <- rep(NA, total.time.steps) # invader population, init.time.steps will stay NA
  
  a.res <- paste('a',sp.resident, sep = '.') # column name of resident interactions
  a.inv <- paste('a',sp.invader, sep = '.') # column name of invader interactions
    
  # running just the resident alone for init.time.steps
  n.res[1] <- N0.r
  for (t in 1:init.time.steps){
    # parameters for environmental conditions--lambda and alpha not correlated
    env.draw.lambda <- sample(c('wet','dry'), size = 1) # env for lambda
    env.draw.alpha <- sample(c('wet','dry'), size = 1) # env for alpha
    
    # resident
    p.res.lambda <- parameters %>% filter(species == sp.resident, treatment == env.draw.lambda)
    p.res.alpha <- parameters %>% filter(species == sp.resident, treatment == env.draw.alpha)
    
    # running one time step using the appropriate environmental parameters
    n.res[t+1] <- pop_equilibrium(N0 = n.res[t], 
                                  s = p.res.lambda$seed.surv, 
                                  g = p.res.lambda$seed.germ, 
                                  lambda = p.res.lambda$lambda, 
                                  a.intra = p.res.alpha[,a.res])
    n.res.impact[t+1] <- n.res[t+1]
   #  n.res.gr[t+1] <- n.res[t+1]/n.res[t]
  }
  
  # running the resident and invader for the remaining time steps
  for (t in (init.time.steps+1):(total.time.steps-1)){
    
    # parameters for environmental conditions--lambda and alpha not correlated
    env.draw.lambda <- sample(c('wet','dry'), size = 1) # env for lambda
    env.draw.alpha <- sample(c('wet','dry'), size = 1) # env for alpha
    
    # resident
    p.res.lambda <- parameters %>% filter(species == sp.resident, treatment == env.draw.lambda)
    p.res.alpha <- parameters %>% filter(species == sp.resident, treatment == env.draw.alpha)
    
    # invader
    p.inv.lambda <- parameters %>% filter(species == sp.invader, treatment == env.draw.lambda)
    p.inv.alpha <- parameters %>% filter(species == sp.invader, treatment == env.draw.alpha)
    
    # invader population
    # N0 of the invader is set to the same value each time
    n.inv[t+1] <- pop_invade(N0 = N0.i, 
                             resident = n.res[t], #** Should be n.res[init.time.steps]
                             s = p.inv.lambda$seed.surv, 
                             g = p.inv.lambda$seed.germ, 
                             lambda = p.inv.lambda$lambda, 
                             a.inter = p.inv.alpha[,a.res])
    
    # resident population as if invader weren't present
    n.res[t+1] <- pop_equilibrium(N0 = n.res[t], 
                                  s = p.res.lambda$seed.surv, 
                                  g = p.res.lambda$seed.germ, 
                                  lambda = p.res.lambda$lambda, 
                                  a.intra = p.res.alpha[,a.res])
    
    # resident population with invader present
    n.res.impact[t+1] <- pop_resident(N0 = N0.i, 
                               resident = n.res[t], 
                               s = p.res.lambda$seed.surv, 
                               g = p.res.lambda$seed.germ, 
                               lambda = p.res.lambda$lambda, 
                               a.intra = p.res.alpha[,a.res],
                               a.inter = p.res.alpha[,a.inv])
    n.res.gr[t+1] <- n.res.impact[t+1]/n.res[t]
  }

  return(tibble(time = 1:total.time.steps, n.inv, n.res, n.res.impact, n.res.gr))
}

```


Plotting function so we can visualize what's going on with the different types of variation
```{r plotting-function}
plot_invasion <- function(df, sp.invader, sp.resident, title){
  df.long <- df %>% 
    select(time, n.inv, n.res.impact) %>% 
    pivot_longer(cols = c(n.res.impact, n.inv),
                 names_to = 'species', values_to = 'n') %>% 
    mutate(n.log = log(n))
  
  ## colors for graphs
  cols <- c(
    as = "#5445b1",
    an = "#f3c483",
    p = "#4DA896",
    lines = "#749dae"
  )
  
  sp.inv.cap <- toupper(sp.invader)
  sp.res.cap <- toupper(sp.resident)
  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    geom_vline(xintercept = time.warm.up, color = cols[['lines']], linetype = 'dashed') +
    theme_cw() +
    scale_color_manual(name = 'Species', values = c(cols[[sp.invader]], cols[[sp.resident]]),
                       labels = c(sp.inv.cap,sp.res.cap)) +
    ggtitle(paste(title, sp.inv.cap, 'invades', sp.res.cap, sep = ' ')) +
    ylim(-2,8.5) +
    xlab('Time') + 
    ylab('Population size (log)')
  
}
```

# Calculating partitions
Function to calculate partitions for a given invader and resident
```{r partition-function}
partition_epsilons <- function(sp.inv, sp.res){
  
  part <- tibble(sp.invader = sp.inv, sp.resident = sp.res)
  
  # no variation
  invade.none <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.i = 1, N0.r = 100, parameters = pars,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               vary.lambda = FALSE, vary.alpha = FALSE) %>%
    filter(time > time.warm.up + 1)
  
  part$inv.e0 <- mean(log(invade.none$n.inv))
  part$res.e0 <- mean(log(invade.none$n.res.gr))
  
  # lambda variation only
  invade.l <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                            N0.i = 1, N0.r = 100, parameters = pars,
                            init.time.steps = time.warm.up, total.time.steps = time.full,
                            vary.lambda = TRUE, vary.alpha = FALSE) %>%
    filter(time > time.warm.up + 1)
  
  part$inv.el <- mean(log(invade.l$n.inv)) - part$inv.e0 # need to check -- maybe should be mean of log
  part$res.el <- mean(log(invade.l$n.res.gr))- part$inv.e0
  
  # alpha variation only
  invade.a <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                            N0.i = 1, N0.r = 100, parameters = pars,
                            init.time.steps = time.warm.up, total.time.steps = time.full,
                            vary.lambda = FALSE, vary.alpha = TRUE) %>%
    filter(time > time.warm.up + 1)
  
  part$inv.ea <- mean(log(invade.a$n.inv)) - part$inv.e0
  part$res.ea <- mean(log(invade.a$n.res.gr)) - part$inv.e0
  
  # full variation
  invade.full <-  run_invasion(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.i = 1, N0.r = 100, parameters = pars,
                               init.time.steps = time.warm.up, total.time.steps = time.full) %>%
    filter(time > time.warm.up + 1)
  
  part$inv.full <- mean(log(invade.full$n.inv)) 
  part$res.full <- mean(log(invade.full$n.res.gr))
  
  part$inv.eint <- part$inv.full - (part$inv.e0 + part$inv.el + part$inv.ea)
  part$res.eint <- part$res.full - (part$res.e0 + part$res.el + part$res.ea)
  
  # no correlation
  invade.no.cov <- run_invasion_no_correlation(sp.invader = sp.inv, sp.resident = sp.res,
                               N0.i = 1, N0.r = 100, parameters = pars,
                               init.time.steps = time.warm.up, total.time.steps = time.full) %>%
    filter(time > time.warm.up + 1)
  
  part$inv.nocov <- mean(log(invade.no.cov$n.inv))
  part$res.nocov <- mean(log(invade.no.cov$n.res.gr))
  
  part$inv.storage <- part$inv.eint - part$inv.nocov
  part$res.storage <- part$res.eint - part$res.nocov
  
  return(part)
}


```

Running that partition function 
```{r partition-run}
time.warm.up <- 50
time.full <- 100

partitions <- partition_epsilons('an','as') %>%
  rbind(partition_epsilons('as','an')) %>%
  pivot_longer(cols = inv.e0:res.storage) %>%
  separate(name, into = c('player', 'partition')) %>%
  pivot_wider(names_from = player, values_from = value) %>%
  mutate(delta = inv - res,
         partition = factor(partition, 
                            levels = c('full','e0','ea','el','eint', 'storage','nocov')))
species_names <- c(
  an = 'AN invades AS',
  as = 'AS invades AN'
)

colors <- c("#5445b1", "#749dae", "#f3c483", "#5c1a33", "#cd3341", "#f7dc6a", "#4DA896")
colors2 <- c("#749dae", rep('darkslategrey', times = 4), rep("#f3c483", times = 2))

ggplot(partitions, aes(x = partition, y = delta, fill = partition)) + 
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_cw() +
  ylab('Partitioning of growth rate when rare') +
  scale_x_discrete(labels = c('full' = 'Full',  'e0' = expression(Delta^0), 'ea' = expression(Delta^alpha),
                              'el' = expression(Delta^lambda), 'eint' = expression(Delta^{alpha*lambda}),
                              'storage' = expression(Delta^(alpha*lambda)), 'nocov' = expression(Delta^{alpha!=lambda})),
                   name = 'Partition') +
  scale_fill_manual(values = colors2) +
  guides(fill = 'none')

# ggsave('partitions.pdf', width = 8, height = 5, units = 'in')

```


# Running and plotting invasions

Full variation (normal run)
```{r normal-run}
time.warm.up <- 50
time.full <- 100
invade.an.full <-  run_invasion(sp.invader = 'an', sp.resident = 'as',
                                N0.i = 1, N0.r = 100, parameters = pars,
                                init.time.steps = time.warm.up, total.time.steps = time.full)

plot_invasion(invade.an.full, 'an', 'as', 'Full variation')


invade.as.full <-  run_invasion(sp.invader = 'as', sp.resident = 'an',
                                N0.i = 1, N0.r = 100, parameters = pars,
                                init.time.steps = time.warm.up, total.time.steps = time.full)

plot_invasion(invade.as.full, 'as', 'an', 'Full variation')

```

No variation
```{r no-variation}
invade.an.none <-  run_invasion(sp.invader = 'an', sp.resident = 'as',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = FALSE, vary.alpha = FALSE)

plot_invasion(invade.an.none, 'an', 'as', 'No variation')

invade.as.none <-  run_invasion(sp.invader = 'as', sp.resident = 'an',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = FALSE, vary.alpha = FALSE)

plot_invasion(invade.as.none, 'as', 'an', 'No variation')
```

Lambda only
```{r lambda-variation}
invade.an.l <-  run_invasion(sp.invader = 'an', sp.resident = 'as',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = TRUE, vary.alpha = FALSE)

plot_invasion(invade.an.l, 'an', 'as', 'Lambda variation')


invade.as.l <-  run_invasion(sp.invader = 'as', sp.resident = 'an',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = TRUE, vary.alpha = FALSE)

plot_invasion(invade.as.l, 'as', 'an', 'Lambda')

```

Alpha only
```{r alpha-variation}
invade.an.a <-  run_invasion(sp.invader = 'an', sp.resident = 'as',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = FALSE, vary.alpha = TRUE)

plot_invasion(invade.an.a, 'an', 'as', 'Alpha variation')

invade.as.a <-  run_invasion(sp.invader = 'as', sp.resident = 'an',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = FALSE, vary.alpha = TRUE)

plot_invasion(invade.as.a, 'as', 'an', 'Alpha variation')

```

Split covariance
```{r no-covariance}
invade.an.no.cov <-  run_invasion_no_correlation(sp.invader = 'an', sp.resident = 'as',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full)

plot_invasion(invade.an.no.cov , 'an', 'as', 'Split covariance')

invade.as.no.cov  <-  run_invasion_no_correlation(sp.invader = 'as', sp.resident = 'an',
                           N0.i = 1, N0.r = 100, parameters = pars,
                           init.time.steps = time.warm.up, total.time.steps = time.full)

plot_invasion(invade.as.no.cov , 'as', 'an', 'Split covariance')

```
