---
title: "MCT-Perennials_partitioning-code"
output: html_document
---
Code used in the manuscript "Fluctuation-dependent coexistence of stage-structured species" to generate figures from the output of the simulation runs. Simulations and partitions were run on a computer cluster and this code is included in the "mct-partitions-functions.R" and "mct-partitions-run.R" files. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)

## ggplot theme
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
    #  strip.text=element_text(size=12),
    #  axis.text=element_text(size=12),
      complete = TRUE
    )
}
```


# Loading in rdata file from partitions run on the computer cluster
```{r partition-plots-load-data}

part.sum.all <- tibble()
for(favorable in c('opposite-consistent', 'same-consistent', 'same-flipped')){
  for(case in c('p', 'as')){
    varying <- 'alpha'
    
    # reading in data
    file.core <- paste0('mct-partitions_', case, '_', varying, '-', favorable)
    load(here(paste0('model-data/', file.core, '.RData')))
    
    # summary of the partition results
    part.long <- simulations %>% map_dfr(bind_rows, .id = "run") %>%
      mutate(partition = factor(partition, 
                                levels = c('full','e0','el','ea','eint',
                                           'storage','nocov')))
    
    part.sum <- part.long %>% 
      group_by(sp.invader, sp.resident, partition, env.ratio) %>%
      dplyr::summarise(
        inv.mean = mean(inv),
        res.mean = mean(res),
        delta.mean = mean(delta),
        inv.sd = sd(inv),
        res.sd = sd(res),
        delta.sd = sd(delta) 
      ) %>% add_column(env.case = sub('-','.',favorable))
    
    part.sum.all <- part.sum.all %>% bind_rows(part.sum)
  }
}


part.sum.all$sp.invader <- factor(part.sum.all$sp.invader, 
                                  levels = c("s.p", "as", "an") )

part.sum.all$sp.resident <- factor(part.sum.all$sp.resident, 
                                  levels = c("s.p", "as", "an") )

part.sum.all$env.case <- factor(part.sum.all$env.case, 
                                  levels = c("opposite.consistent", "same.flipped", "same.consistent") )

# graph x-axes by proportion of env intrinsically favorable to our focal species
# (perennial or seedbanking annual) -- in the original parameter files
# these were the ones that switched with the different scenarios
part.sum.all <- part.sum.all %>% 
  mutate(
    proportion.favorable = ifelse(env.case == "opposite.consistent", 1 - part.sum.all$env.ratio, part.sum.all$env.ratio)
  ) 

```


# Plotting figures

Shared components for all plots
```{r partition-plots-components}
partition.legend <- c(
  'full' = 'Full',  
  'e0' = expression(Delta^0), 
  'el' = expression(Delta^lambda), 
  'ea' = expression(Delta^alpha),
  'eint' = expression(Delta^{alpha*lambda}),
  'storage' = expression(Delta^(alpha*lambda)), 
  'nocov' = expression(Delta^{paste(alpha, "#", lambda)})
)


cb_palette <- c("black", "#009E73", "#56B4E9", "#E69F00", "darkorchid4")
cb_palette_long <- c("black", "#009E73", "#56B4E9", "#E69F00", "darkorchid4",
                            "#D55E00", "#0072B2")
cb_palette_int <- c("black",  "darkorchid4", "#D55E00", "#0072B2")

```

Figure 2: Across environmental conditions, perennial and seedbanking annual as the invaders, non-seedbanking annual as the resident
```{r partition-plots-figure-2}

part.sum.inv <- part.sum.all %>% 
  filter(sp.resident == 'an', 
         partition %in% c('full', 'e0', 'el', 'ea', 'eint'))

environmental_condition <- as_labeller(c(
  opposite.consistent = "Opposite env favorable",
  same.consistent = "Same env favorable \n competition stronger",
  same.flipped = "Same env favorable \n competition weaker",
  s.p = "Invader: Perennial \n Resident: Nonseedbanking Annual",
  as = "Invader: Seedbanking Annual \n Resident: Nonseedbanking Annual"
))

part.sum.inv %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(env.case), cols = vars(sp.invader), scales = "free",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = delta.mean - delta.sd, 
                               ymax = delta.mean + delta.sd,
                               fill = partition)) +
  geom_line(aes(y = delta.mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.05, 0.93),
        legend.text.align = 0) +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate')

# ggsave('mct-partitions_figure-2.pdf', width = 8.5, height = 8.5, units = 'in')
```


Figure 3: across environmental conditions, perennial and seedbanking annual as the resident, non-seedbanking annual as the invader
```{r partition-plots-figure-3}

part.sum.res <- part.sum.all %>% 
  filter(sp.invader == 'an', 
         partition %in% c('full', 'e0', 'el', 'ea', 'eint'))

environmental_condition <- as_labeller(c(
  opposite.consistent = "Opposite env favorable",
  same.consistent = "Same env favorable \n competition stronger",
  same.flipped = "Same env favorable \n competition weaker",
  s.p = "Invader: Nonseedbanking Annual \n Resident: Perennial",
  as = "Invader: Nonseedbanking Annual \n Resident: Seedbanking Annual"
))

part.sum.res %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(env.case), cols = vars(sp.resident), scales = "free",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = delta.mean - delta.sd, 
                               ymax = delta.mean + delta.sd,
                               fill = partition)) +
  geom_line(aes(y = delta.mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.93, 0.93),
        legend.text.align = 0) +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = '') +
  xlab('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate')

# ggsave('mct-partitions_figure-3.pdf', width = 8.5, height = 8.5, units = 'in')
```

Supplemental figures
```{r supplemental-figure}
# Supplemental figure matching Figure 2
# With the additional partitions of the storage effect
 
 part.sum.supp.2 <- part.sum.all %>% 
  filter(sp.resident == 'an',
         partition %in% c('full', 'eint', 'storage', 'nocov'))

environmental_condition <- as_labeller(c(
  opposite.consistent = "Opposite env favorable",
  same.consistent = "Same env favorable \n competition stronger",
  same.flipped = "Same env favorable \n competition weaker",
  s.p = "Invader: Perennial \n Resident: Nonseedbanking Annual",
  as = "Invader: Seedbanking Annual \n Resident: Nonseedbanking Annual"
))


part.sum.supp.2 %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(env.case), cols = vars(sp.invader), scales = "free",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = delta.mean - delta.sd, 
                               ymax = delta.mean + delta.sd,
                               fill = partition)) +
  geom_line(aes(y = delta.mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.05, 0.9),
        legend.text.align = 0) +
  scale_color_manual(values = cb_palette_int, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette_int, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate')

# ggsave('mct-partitions_figure-supp-2.pdf', width = 8.5, height = 8.5, units = 'in')
 

# Supplemental figure matching Figure 3
# With the additional partitions of the storage effect
 
 part.sum.supp.3 <- part.sum.all %>% 
  filter(sp.invader == 'an',
         partition %in% c('full', 'eint', 'storage', 'nocov'))
 
 environmental_condition <- as_labeller(c(
  opposite.consistent = "Opposite env favorable",
  same.consistent = "Same env favorable \n competition stronger",
  same.flipped = "Same env favorable \n competition weaker",
  s.p = "Invader: Nonseedbanking Annual \n Resident: Perennial",
  as = "Invader: Nonseedbanking Annual \n Resident: Seedbanking Annual"
))

part.sum.supp.3 %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(env.case), cols = vars(sp.resident), scales = "free",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = delta.mean - delta.sd, 
                               ymax = delta.mean + delta.sd,
                               fill = partition)) +
  geom_line(aes(y = delta.mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.45, 0.93),
        legend.text.align = 0) +
  scale_color_manual(values = cb_palette_int, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette_int, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate')

# ggsave('mct-partitions_figure-supp-3.pdf', width = 8.5, height = 8.5, units = 'in')
```


Explicitly depicting the Delta values as the Invader-Resident Comparison

Corresponding to Figure 2, perennial or seedbanking annual invader and non-seedbanking annual resident. 
```{r invader-resident}
part.sum.inv.long <- part.sum.inv %>% 
  pivot_longer(cols = c('delta.mean', 'inv.mean', 'res.mean', 'delta.sd', 'inv.sd', 'res.sd'),
               names_to = 'names', values_to = 'value') %>%
  separate_wider_delim(cols = 'names', delim = '.', names = c('epsilon', 'metric')) %>%
  pivot_wider(names_from = metric, values_from = value)

environmental_condition <- as_labeller(c(
  s.p = "Invader: Perennial \n Resident: Nonseedbanking Annual",
  as = "Invader: Seedbanking Annual \n Resident: Nonseedbanking Annual",
  delta = "Delta",
  inv = "Invader",
  res = "Resident"
))

  
# Same environment favorable, competition weaker in favorable environment
part.sum.inv.same.flipped <- part.sum.inv.long %>% filter(env.case == 'same.flipped')

part.sum.inv.same.flipped %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(epsilon), cols = vars(sp.invader), scales = "fixed",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = mean - sd, 
                               ymax = mean + sd,
                               fill = partition)) +
  geom_line(aes(y = mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.05, 0.95),
        legend.text.align = 0, 
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.7, "lines")) +
  guides(shape = guide_legend(override.aes = list(size = 0.9)),
         color = guide_legend(override.aes = list(size = 0.9))) +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate') +
  ggtitle('Scenario 2: Same Environment Favorable; Competition Weaker')

 ggsave('mct-partitions_figure-supp_dir-2a.pdf', width = 8.5, height = 8.5, units = 'in')

# Same environment favorable, competition stronger in favorable environment
part.sum.inv.same.consistent <- part.sum.inv.long %>% filter(env.case == 'same.consistent')

part.sum.inv.same.consistent %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(epsilon), cols = vars(sp.invader), scales = "fixed",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = mean - sd, 
                               ymax = mean + sd,
                               fill = partition)) +
  geom_line(aes(y = mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.05, 0.95),
        legend.text.align = 0, 
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.7, "lines")) +
  guides(shape = guide_legend(override.aes = list(size = 0.9)),
         color = guide_legend(override.aes = list(size = 0.9))) +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate') +
  ggtitle('Scenario 3: Same Environment Favorable; Competition Stronger')

 ggsave('mct-partitions_figure-supp_dir-2b.pdf', width = 8.5, height = 8.5, units = 'in')


```

Corresponding to Figure 3, perennial or seedbanking annual resident and non-seedbanking annual invader. 
```{r invader-resident-2}
part.sum.res.long <- part.sum.res %>% 
  pivot_longer(cols = c('delta.mean', 'inv.mean', 'res.mean', 'delta.sd', 'inv.sd', 'res.sd'),
               names_to = 'names', values_to = 'value') %>%
  separate_wider_delim(cols = 'names', delim = '.', names = c('epsilon', 'metric')) %>%
  pivot_wider(names_from = metric, values_from = value)

environmental_condition <- as_labeller(c(
  s.p = "Invader: Nonseedbanking Annual \n Resident: Perennial",
  as = "Invader: Nonseedbanking Annual \n Resident: Seedbanking Annual",
  delta = "Delta",
  inv = "Invader",
  res = "Resident"
))

# smaller legend size
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
  myPlot +
    guides(shape = guide_legend(override.aes = list(size = pointSize)),
           color = guide_legend(override.aes = list(size = pointSize))) +
    theme(legend.title = element_text(size = textSize), 
          legend.text  = element_text(size = textSize),
          legend.key.size = unit(spaceLegend, "lines"))
}

  
# Same environment favorable, competition weaker in favorable environment
part.sum.res.same.flipped <- part.sum.res.long %>% filter(env.case == 'same.flipped')

part.sum.res.same.flipped %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(epsilon), cols = vars(sp.resident), scales = "free",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = mean - sd, 
                               ymax = mean + sd,
                               fill = partition)) +
  geom_line(aes(y = mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.05, 0.95),
        legend.text.align = 0, 
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.7, "lines")) +
  guides(shape = guide_legend(override.aes = list(size = 0.9)),
         color = guide_legend(override.aes = list(size = 0.9))) +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate') +
  ggtitle('Scenario 2: Same Environment Favorable; Competition Weaker')

 ggsave('mct-partitions_figure-supp_dir-3a.pdf', width = 8.5, height = 8.5, units = 'in')

# Same environment favorable, competition stronger in favorable environment
part.sum.res.same.consistent <- part.sum.res.long %>% filter(env.case == 'same.consistent')

part.sum.res.same.consistent %>% ggplot(aes(x = proportion.favorable)) +
  facet_grid(rows = vars(epsilon), cols = vars(sp.resident), scales = "free",
             labeller = environmental_condition) +
  geom_ribbon(alpha = 0.3, aes(ymin = mean - sd, 
                               ymax = mean + sd,
                               fill = partition)) +
  geom_line(aes(y = mean, color = partition)) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.05, 0.95),
        legend.text.align = 0, 
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.7, "lines")) +
  guides(shape = guide_legend(override.aes = list(size = 0.9)),
         color = guide_legend(override.aes = list(size = 0.9))) +
  scale_color_manual(values = cb_palette, labels = partition.legend, name = '') +
  scale_fill_manual(values = cb_palette, labels = partition.legend, name = '') +
  xlab ('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Partitioning of low density growth rate') +
  ggtitle('Scenario 3: Same Environment Favorable; Competition Stronger')

 ggsave('mct-partitions_figure-supp_dir-3b.pdf', width = 8.5, height = 8.5, units = 'in')

```

