---
title: "MCT-Perennials_partitioning-code_new"
output: html_document
---
A simplified version of the code used in the manuscript "Fluctuation-dependent coexistence of stage-structured species" to demonstrate how the partitioning works with a smaller number of simulation runs. The actual code was run on a computer cluster and is included in the "mct-partitions-functions.R" and "mct-partitions-run.R" files. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)

## ggplot theme
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      #  strip.text=element_text(size=12),
      #  axis.text=element_text(size=12),
      complete = TRUE
    )
}
```

## Parameters and set up

This step defines which coexistence pair we're looking at, seedbanking annual with non-seedbanking annual, or perennial with non-seedbanking annual. Then sets parameters we're going to use for alpha and lambda under different environmental conditions. 
```{r parameters}

# seedbanking annual - nonseedbanking annual parameters
pars.as.an <- read.csv(here('model-parameters/model-parameters_as_alpha-same-flipped.csv')) %>%
  mutate(stem.surv = NULL) 

# perennial - nonseedbanking annual parameters
pars.p.an <- read.csv(here('model-parameters/model-parameters_p_alpha-same-flipped.csv')) %>%
  mutate(seed.surv = NULL)

```


## Functions for species interactions and simulation runs
Functions for one step of species interaction calculations. Works whether it's both species present or just one--if it's one, the other population is just at 0. Takes in the full parameter df for all the species (just for one environmental condition though) and a list of species population sizes. If it's just one species running to equilibrium then it's N0 will be 0. 
```{r species-interactions}
pop_interactions <- function(N0, param){
  
  # seedbanking annual - nonseedbanking annual
  if(pair == 'as.an'){
    pars.as <- param %>% filter(species == 'as')
    pars.an <- param %>% filter(species == 'an')
    
    n.as <- N0$as * (1 - pars.as$seed.germ) * pars.as$seed.surv + 
      N0$as * pars.as$seed.germ *pars.as$lambda /
      (1 + pars.as$a.an * pars.an$seed.germ * N0$an +
         pars.as$a.as * pars.as$seed.germ * N0$as)
    
    n.an <- N0$an * pars.an$seed.germ *pars.an$lambda /
      (1 + pars.an$a.an * pars.an$seed.germ * N0$an +
         pars.an$a.as * pars.as$seed.germ * N0$as)
    
    n.all <- tibble(an = n.an,
                    as = n.as)
  }
  
  # perennial - nonseedbanking annual
  if(pair == 'p.an'){
    pars.s <- param %>% filter(species == 's')
    pars.p <- param %>% filter(species == 'p')
    pars.an <- param %>% filter(species == 'an')
    
    n.s <- N0$p*pars.p$lambda /
      (1 + pars.p$a.an * pars.an$seed.germ * N0$an +
         pars.p$a.s * pars.s$seed.germ * N0$s +
         pars.p$a.p * N0$p)
    
    n.p <- N0$p * pars.p$stem.surv +
      (N0$s * pars.s$seed.germ * pars.s$stem.surv) /
      (1 + pars.s$a.an * pars.an$seed.germ * N0$an +
         pars.s$a.s * pars.s$seed.germ * N0$s +
         pars.s$a.p * N0$p)
    
    n.an <- N0$an * pars.an$seed.germ *pars.an$lambda /
      (1+ pars.an$a.an * pars.an$seed.germ * N0$an +
         pars.an$a.s * pars.s$seed.germ * N0$s +
         pars.an$a.p * N0$p)
    
    n.all <- tibble(an = n.an,
                    s = n.s,
                    p = n.p)
  }
  return(n.all)
}

```


Function to run one time series, first with sp.resident alone for init.time.steps alone starting from population size N0.r. Then the invader species is added in at each time step at low density from init.time.steps to total.time.steps. Parameter names sp.invader and sp.resident can be either a single value (ex: 'a') or multiple values (ex: c('s','p')) to accommodate stage-structured species.

This function runs with all parameters varying. All the column names are based on the actual species names.
```{r running-full}
perennial_distribution <- function(change.limit = 0.01, n.old = N0.together, 
                                   p.env = p.env, sp.resident, sp.invader) {
  n.old.inv <- n.old[sp.invader]
  change <- 1
  
  while(change > change.limit){
    n.update <- pop_interactions(N0 = n.old, param = p.env) # run one time step
    
    n.per.total <- sum(n.update[sp.invader]) # total perennial population
    n.update[sp.invader] <- n.update[sp.invader]/n.per.total # normalize and update perennial population
    n.update[sp.resident] <- n.old[sp.resident] # keep resident population the same (do not update)
    
    # calculate change in perennial population distribution
    n.new.inv <- n.update[sp.invader]
    change <- abs(n.old.inv[1] - n.new.inv[1])
    
    # reset values for next loop
    n.old <- n.update
    n.old.inv <- n.new.inv[1]
  }
  return(n.update) # return dataframe with all population size values
}


# First run with all parameters varying. This is where stage structure distribution is calculated
# and this what is used for all the following non-varying runs
run_full <- function(sp.invader, sp.resident, N0.r, parameters, env.cond,
                     init.time.steps = 50, total.time.steps = 100) {
  
  # populations with resident species not impacted by the invader
  if(pair == 'as.an'){
    n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                    as = as.numeric(rep(NA, total.time.steps)))
  } 
  else if(pair == 'p.an'){
    n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                    s = as.numeric(rep(NA, total.time.steps)),
                    p = as.numeric(rep(NA, total.time.steps)))
  }
  
  # setting initial populations
  n.all[1,sp.resident] <- N0.r
  n.all[1,sp.invader] <- 0
  
  # name of column(s) for resident species pop when impacted by invading pop
  sp.res.impact <- paste(sp.resident, 'impact', sep = '.')
  # names of column(s) for invader stage distribution
  sp.inv.dist <- paste(sp.invader, 'dist', sep = '.')
  
  for(t in 1:init.time.steps){
    # parameters all varying
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
    # running one time step with the appropriate environmental parameters
    n.all[t+1,] <- pop_interactions(N0 = n.all[t,], param = p.env)
    
  }
  
  # figuring out rough invader population distribution if the invader is perennial
  # using average parameters and resident population size during the final warm-up step
  # makes it faster for running in for loop
  if(length(sp.invader) > 1) {
    
    n.inv.start <- 1/length(sp.invader)
    n.old <- n.all[init.time.steps,]
    n.old[sp.invader] <- n.inv.start
    
    # while-loop with updates happens in perennial_distribution function
    n.old <- perennial_distribution(change.limit = 0.1, n.old = n.old, p.env = p.env,
                                    sp.resident, sp.invader) 
  }
  
  # running the resident and invader for the remaining time steps
  for (t in (init.time.steps+1):(total.time.steps-1)){
    # parameters for the varying env condition
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
    # running invader and resident together
    N0.together <- n.all[t,]
    
    # figuring out starting invader population distribution if the invader is perennial
    # starting from the n.old found for average conditions above
    if(length(sp.invader) > 1){
      n.old[sp.resident] <- n.all[t,sp.resident]
      n.update <- perennial_distribution(change.limit = 0.01, n.old = n.old, p.env = p.env,
                                         sp.resident, sp.invader) # function
      
      N0.together[sp.invader] <- n.update[sp.invader] # using the equilibrium stage distribution from above
    } else {
      N0.together[sp.invader] <- 1 # if the invader has only one life stage
    }
    n.all[t,sp.inv.dist] <- N0.together[sp.invader]
    
    
    # running one time step with the appropriate environmental parameters
    n.res.impact <- pop_interactions(N0 = N0.together, param = p.env)
    
    # saving this output as the invader population and the impacted resident population
    n.all[t+1,sp.invader] <- n.res.impact[sp.invader]
    n.all[t+1,sp.res.impact] <- n.res.impact[sp.resident]
    
    # a version of this with the resident unaffected by the invader
    N0.res.only <- n.all[t,] %>% select(all_of(sp.invader), all_of(sp.resident))
    N0.res.only[sp.invader] <- 0
    n.res.only <- pop_interactions(N0 = N0.res.only, param = p.env)
    n.all[t+1,sp.resident] <- n.res.only[sp.resident]
    
    # growth rate of the resident
    sp.res.gr <- paste('gr', paste(sp.resident, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.res.gr] <- sum(n.all[t+1,sp.res.impact])/sum(n.all[t,sp.resident])
    
    # growth rate of the invader
    sp.inv.gr <- paste('gr', paste(sp.invader, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.inv.gr] <- sum(n.all[t+1,sp.invader])
    
  }
  
  n.all$time <- 1:total.time.steps
  return(n.all)
  
}

```



Function to calculate growth rates for different combinations of varying parameters. For partitions where lambda is set to not vary, also sets germination and survival terms to not vary. All the column names are based on the actual species names. The next section of the code can use this to calculate growth rates and average across the time series. 

```{r running-invasion}

# similar but modified function for the following runs with different varying parameters
# this uses the output from the full run above
run_varying <- function(sp.invader, sp.resident, parameters, env.cond,
                        vary.lambda = TRUE, vary.alpha = TRUE, covary = TRUE,
                        init.time.steps = 50, total.time.steps = 100, n.full) {
  
  
  # populations with resident species not impacted by the invader
  if(pair == 'as.an'){
    n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                    as = as.numeric(rep(NA, total.time.steps)))
  } 
  else if(pair == 'p.an'){
    n.all <- tibble(an = as.numeric(rep(NA, total.time.steps)),
                    s = as.numeric(rep(NA, total.time.steps)),
                    p = as.numeric(rep(NA, total.time.steps)))
  }
  
  
  # name of column(s) for resident species pop when impacted by invading pop
  sp.res.impact <- paste(sp.resident, 'impact', sep = '.')
  # names of column(s) for invader stage distribution
  sp.inv.dist <- paste(sp.invader, 'dist', sep = '.')
  
  # parameters for the average env condition
  p.ave <- parameters %>% filter(treatment == 'average')
  
  # columns used for the non-varying parameters
  alpha.columns <- c(paste('a', sp.invader, sep = '.'), 
                     paste('a', sp.resident, sep = '.'))
  lambda.columns <- parameters %>% 
    select(!c(all_of(alpha.columns),'species','treatment')) %>% 
    names()
  
  # # environmental condition vectors for the no covariance option 
  env.lambda <- sample(env.cond, size = length(env.cond), replace = FALSE)
  env.alpha <- sample(env.cond, size = length(env.cond), replace = FALSE)
  
  
  # running the resident and invader for the remaining time steps
  for (t in (init.time.steps+1):(total.time.steps-1)){
    
    
    # parameters for the varying env condition
    p.env <- parameters %>% filter(treatment == env.cond[t])
    
    # adjusting non-varying or non-covarying parameters
    
    # constant lambda
    if(vary.lambda == FALSE){
      p.env[,lambda.columns] <- p.ave[,lambda.columns]
    }
    
    # constant alpha
    if(vary.alpha == FALSE){
      p.env[,alpha.columns] <- p.ave[,alpha.columns]
    }
    
    # parameters for no covariance between alpha and lambda 
    if(covary == FALSE){
      lambda.env  <- parameters %>% filter(treatment == env.lambda[t])
      alpha.env <- parameters %>% filter(treatment == env.alpha[t])
      p.env[,lambda.columns] <- lambda.env[,lambda.columns]
      p.env[,alpha.columns] <- alpha.env[,alpha.columns]
    }
    
    
    # Starting population sizes are based on the fully varying simulation only
    # Resident starting population size is un-impacted (without the invader present)
    # Invader total starting population size is 1, stage distribution is the same
    # stage distribution from the fully varying run
    
    N0.together <- n.full[t,] 
    if(length(sp.invader) > 1){
      N0.together[sp.invader] <- N0.together[sp.inv.dist] 
    } else {
      N0.together[sp.invader] <- 1 # if the invader has only one life stage
    }
    
    
    # running one time step with the appropriate environmental parameters
    n.res.impact <- pop_interactions(N0 = N0.together, param = p.env)
    
    # saving this output as the invader population and the impacted resident population
    n.all[t+1,sp.invader] <- n.res.impact[sp.invader]
    n.all[t+1,sp.res.impact] <- n.res.impact[sp.resident]
    
    # growth rate of the resident from full variation starting point
    sp.res.gr <- paste('gr', paste(sp.resident, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.res.gr] <- sum(n.all[t+1,sp.res.impact])/sum(N0.together[sp.resident])
    
    # growth rate of the invader (always from an initial total population of 1)
    sp.inv.gr <- paste('gr', paste(sp.invader, collapse = '.'), sep = '.') # column name
    n.all[t+1,sp.inv.gr] <- sum(n.all[t+1,sp.invader])
    
  }
  
  n.all$time <- 1:total.time.steps
  return(n.all) 
  
}

```

Basic plotting functions to visualize the variation through time
```{r plotting-function}
plot_invasion <- function(df){ # plots actual population sizes
  df.long <- df %>% 
    pivot_longer(cols = !c(time, starts_with('gr')),
                 names_to = 'species', values_to = 'n',
                 values_drop_na = TRUE) %>% 
    mutate(n.log = log(n))
  
  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    theme_cw() +
    xlab('Time') + 
    ylab('Population size (log)')
  
}

plot_gr <- function(df) { # plots population growth rates
  df.long <- df %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', values_to = 'n',
                 values_drop_na = TRUE) %>% 
    mutate(n.log = log(n))
  
  
  ggplot(df.long, aes(x = time, y = n.log, color = species)) +
    geom_line() +
    geom_point() +
    theme_cw() +
    xlab('Time') + 
    ylab('Growth Rate (log)')
}
```

## Running and plotting invasions 

Set up reading in parameters, used in all the following sections of code
```{r parameters}
# CHOOSE: This determines which coexistence pair we're looking at
case <- 'as'
#case <- 'p'

pair <- paste(case, 'an', sep = '.')

varying <- 'alpha'

 favorable <- 'same-consistent'
# favorable <- 'opposite-consistent'
# favorable <- 'same-flipped'


# file names
file.core <- paste0(case, '_', varying, '-', favorable)
main.wd <- './'

file.in <- paste0(main.wd, 'model-parameters/model-parameters_', file.core, '.csv')

file.out <- paste0(main.wd, 'model-data/mct-partitions_', file.core, '.RData')

# SET PARAMETER VALUES 
pars <- read.csv(file.in) 


# select the relevant parameter set
if(pair == 'as.an') {
  sp.1 <- 'an'
  sp.2 <- 'as'
  sp.order <- c('an','as')
  pars <- pars %>% mutate(stem.surv = NULL) 
}
if(pair == 'p.an'){
  sp.1 <- 'an'
  sp.2 <- c('s','p')
  sp.order <- c('an','s','p')
  pars <- pars %>% mutate(seed.surv = NULL)
}

```


Full variation with no partitions, single run
```{r normal-run}
time.warm.up <- 100
time.full <- 200


# two equally probable environmental conditions
env.ratio <- 0.5
env.draw <- rbinom(time.full, 1, env.ratio) 
env.condition <- if_else(env.draw > 0, 'wet','dry')

pars.full <- pars %>% mutate(weight = ifelse(treatment == 'wet', env.ratio, 1-env.ratio))

pars.ave <- pars.full %>% 
  group_by(species) %>%
  summarise_if(is.numeric,
               ~ weighted.mean(., weight)) %>%
  mutate(weight = NA, treatment = 'average') %>% 
  arrange(factor(species, levels = sp.order)) # need to keep them in the right order for later

pars.full <- pars.full %>% full_join(pars.ave)

invade.1.full <-  run_full(sp.invader = sp.1, sp.resident = sp.2, N0.r = 100, 
                           parameters = pars.full, env.cond = env.condition,
                           # vary.lambda = TRUE, vary.alpha = TRUE, 
                           init.time.steps = time.warm.up, total.time.steps = time.full) 


invade.2.full <-  run_full(sp.invader = sp.2, sp.resident = sp.1,
                           N0.r = 100, parameters = pars.full, env.cond = env.condition,
                           # vary.lambda = TRUE, vary.alpha = TRUE,
                           init.time.steps = time.warm.up, total.time.steps = time.full)


plot_invasion(filter(invade.1.full, time > time.warm.up))
plot_gr(invade.1.full)


plot_invasion(filter(invade.2.full, time > time.warm.up))
plot_gr(invade.2.full)

```

Testing the non-varying runs (relies on code in section above). 
```{r variable-run}


# check that when everything is varying as with the full variation the results are consistent
invade.1.full.test <- run_varying(sp.invader = sp.1, sp.resident = sp.2, 
                                  parameters = pars.full, env.cond = env.condition,
                                  vary.lambda = TRUE, vary.alpha = TRUE, covary = TRUE,
                                  init.time.steps = time.warm.up, total.time.steps = time.full,
                                  n.full = invade.1.full)

invade.2.full.test <- run_varying(sp.invader = sp.2, sp.resident = sp.1, 
                                  parameters = pars.full, env.cond = env.condition,
                                  vary.lambda = TRUE, vary.alpha = TRUE, covary = TRUE,
                                  init.time.steps = time.warm.up, total.time.steps = time.full,
                                  n.full = invade.2.full)


# now doing the different variations
invade.1.lambda <- run_varying(sp.invader = sp.1, sp.resident = sp.2, 
                               parameters = pars.full, env.cond = env.condition,
                               vary.lambda = FALSE, vary.alpha = TRUE, covary = TRUE,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               n.full = invade.1.full)

invade.2.lambda <- run_varying(sp.invader = sp.2, sp.resident = sp.1, 
                               parameters = pars.full, env.cond = env.condition,
                               vary.lambda = FALSE, vary.alpha = TRUE, covary = TRUE,
                               init.time.steps = time.warm.up, total.time.steps = time.full,
                               n.full = invade.2.full)


```


## Calculating partitions 
Function to calculate partitions for a given invader and resident
```{r partition-function}
partition_epsilons <- function(sp.inv, sp.res, pars, 
                               env.condition, time.warm.up, time.full){
  res.name <- paste(sp.res, collapse = '.')
  inv.name <- paste(sp.inv, collapse = '.')
  
  part <- tibble(sp.invader = inv.name, sp.resident = res.name)
  
  # full variation
  invade.full <-  run_full(sp.invader = sp.inv, sp.resident = sp.res,
                           N0.r = 100, parameters = pars, env.cond = env.condition,
                           init.time.steps = time.warm.up, total.time.steps = time.full) 
  invade.full.2 <- invade.full %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  part$inv.full <- invade.full.2 %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  part$res.full <- invade.full.2 %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  # no variation full run
  invade.none <-  run_varying(sp.invader = sp.inv, sp.resident = sp.res,
                              parameters = pars, env.cond = env.condition,
                              init.time.steps = time.warm.up, 
                              total.time.steps = time.full,
                              vary.lambda = FALSE, vary.alpha = FALSE, covary = TRUE,
                              n.full = invade.full) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  part$inv.e0 <- invade.none %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  part$res.e0 <- invade.none %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  # lambda variation only
  invade.l <- run_varying(sp.invader = sp.inv, sp.resident = sp.res,
                          parameters = pars, env.cond = env.condition,
                          init.time.steps = time.warm.up, total.time.steps = time.full,
                          vary.lambda = TRUE, vary.alpha = FALSE, covary = TRUE,
                          n.full = invade.full) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  l.inv.mean <- invade.l %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  l.res.mean <- invade.l %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.el <- l.inv.mean - part$inv.e0 
  part$res.el <- l.res.mean- part$res.e0
  
  # alpha variation only
  invade.a <-  run_varying(sp.invader = sp.inv, sp.resident = sp.res,
                           parameters = pars, env.cond = env.condition,
                           init.time.steps = time.warm.up, total.time.steps = time.full,
                           vary.lambda = FALSE, vary.alpha = TRUE, covary = TRUE,
                           n.full = invade.full) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  a.inv.mean <- invade.a %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  a.res.mean <- invade.a %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  part$inv.ea <- a.inv.mean - part$inv.e0 
  part$res.ea <- a.res.mean- part$res.e0
  
  part$inv.eint <- part$inv.full - (part$inv.e0 + part$inv.el + part$inv.ea)
  part$res.eint <- part$res.full - (part$res.e0 + part$res.el + part$res.ea)
  
  # no correlation
  invade.nocov <- run_varying(sp.invader = sp.inv, sp.resident = sp.res,
                              parameters = pars, env.cond = env.condition,
                              init.time.steps = time.warm.up, total.time.steps = time.full,
                              vary.lambda = TRUE, vary.alpha = TRUE, covary = FALSE,
                              n.full = invade.full) %>%
    filter(time > time.warm.up + 1) %>% 
    pivot_longer(cols = starts_with('gr'),
                 names_to = 'species', names_prefix = 'gr.',
                 values_to = 'growth.rate',
                 values_drop_na = TRUE) %>% 
    mutate(gr.log = log(growth.rate))
  
  nocov.inv <- invade.nocov %>% 
    filter(species == inv.name) %>%
    select(gr.log) %>% colMeans()
  nocov.res <- invade.nocov %>% 
    filter(species == res.name) %>%
    select(gr.log) %>% colMeans()
  
  
  part$inv.nocov <- nocov.inv - (part$inv.e0 + part$inv.el + part$inv.ea)
  part$res.nocov <- nocov.res - (part$res.e0 + part$res.el + part$res.ea)
  
  # storage effect
  part$inv.storage <- part$inv.eint - part$inv.nocov
  part$res.storage <- part$res.eint - part$res.nocov
  
  return(part)
}


```

Running that partition function 
```{r partition-run}

env.ratio <- 0.5 # ratio of wet years to dry years

# adding rows for the weighted average of the parameters
pars.full <- pars %>% mutate(weight = ifelse(treatment == 'wet', env.ratio, 1-env.ratio))

pars.ave <- pars.full %>% 
  group_by(species) %>%
  summarise_if(is.numeric,
               ~ weighted.mean(., weight)) %>%
  mutate(weight = NA, treatment = 'average') %>% 
  arrange(factor(species, levels = sp.order)) # need to keep them in the right order for later

pars.full <- pars.full %>% full_join(pars.ave)

time.warm.up <- 100
time.full <- 200


# Environmental conditions
env.draw <- rbinom(time.full, 1, env.ratio) # sequence of the environment 
env.condition <- if_else(env.draw > 0, 'wet','dry')


# testing partition_epsilons function
test.1 <- partition_epsilons(sp.inv = sp.2, sp.res = sp.1, pars.full, 
                             env.condition, time.warm.up, time.full)
test.2 <- partition_epsilons(sp.inv = sp.1, sp.res = sp.2, pars.full, 
                             env.condition, time.warm.up, time.full)


partitions <- partition_epsilons(sp.inv = sp.2, sp.res = sp.1, pars.full, 
                                 env.condition, time.warm.up, time.full) %>%
  rbind(partition_epsilons(sp.inv = sp.1, sp.res = sp.2, pars.full, 
                           env.condition, time.warm.up, time.full)) %>%
  #pivot_longer(cols = inv.e0:res.eint) %>%
  pivot_longer(cols = inv.full:res.storage) %>%
  separate(name, into = c('player', 'partition')) %>%
  pivot_wider(names_from = player, values_from = value) %>%
  mutate(delta = inv - res,
         partition = factor(partition, 
                            levels = c('full','e0','el','ea','eint',
                                       'storage','nocov')))

# Titles for invasion interactions
if(pair == 'as.an'){
  species_names <- c(
    an = 'AN invades AS',
    as = 'AS invades AN'
  )
  file.out <- 'partitions-AS-AN.pdf'
}
if(pair == 'p.an'){
  species_names <- c(
    an = 'AN invades P',
    s.p = 'P invades AN'
  )
  file.out <- 'partitions-P-AN.pdf'
}


# Visualizing output of the partition for the single run
cb_palette_long <- c("black", "#009E73", "#56B4E9", "#E69F00", "darkorchid4",
                            "#D55E00", "#0072B2")
                            
ggplot(partitions, aes(x = partition, y = delta, fill = partition)) + 
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_cw() +
  ylab('Partitioning Delta terms') +
  scale_x_discrete(labels = c('full' = 'Full',  'e0' = expression(Delta^0),
                              'el' = expression(Delta^lambda), 'ea' = expression(Delta^alpha),
                              'eint' = expression(Delta^{alpha*lambda}),
                              'storage' = expression(Delta^(alpha*lambda)), 
                              'nocov' = expression(Delta^{alpha!=lambda})),
                   name = 'Partition') +
  scale_fill_manual(values = cb_palette_long) +
  guides(fill = 'none')


# Delta as invader-resident comparison
partitions.long <- partitions %>% 
  pivot_longer(cols = c('delta', 'inv', 'res'), names_to = 'epsilon', values_to = 'value')

ggplot(partitions.long, aes(x = partition, y = value, fill = partition)) + 
  facet_grid(cols = vars(sp.invader), rows = vars(epsilon),
             labeller = labeller(sp.invader = species_names)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_cw() +
  ylab('Partitioning Delta terms') +
  scale_x_discrete(labels = c('full' = 'Full',  'e0' = expression(Delta^0),
                              'el' = expression(Delta^lambda), 'ea' = expression(Delta^alpha),
                              'eint' = expression(Delta^{alpha*lambda}),
                              'storage' = expression(Delta^(alpha*lambda)), 
                              'nocov' = expression(Delta^{alpha!=lambda})),
                   name = 'Partition') +
  scale_fill_manual(values = cb_palette_long) +
  guides(fill = 'none')


```

Running over an environmental sequence instead of a particular env.ratio. This was used for testing before running on the computer cluster (can take a while to run depending on settings)


```{r env-seq}


# number of time steps for each run
time.warm.up <- 100
time.full <- 200

# initialize dataframe
partitions.env <- tibble(sp.invader = character(), sp.resident = character(),
                         partition = character(),
                         inv = numeric(), res = numeric(), delta = numeric(), 
                         env.ratio = numeric())
env.seq <- seq(0, 1, by = 0.1) # ratio of wet years to dry years

for(env.ratio in env.seq){
  # adding rows for the weighted average of the parameters
  pars.full <- pars %>% mutate(weight = ifelse(treatment == 'wet', env.ratio, 1 - env.ratio))
  
  pars.ave <- pars.full %>% 
    group_by(species) %>%
    summarise_if(is.numeric,
                 ~ weighted.mean(., weight)) %>%
    mutate(weight = NA, treatment = 'average') %>% 
    arrange(factor(species, levels = sp.order)) # need to keep them in the right order for later
  
  pars.full <- pars.full %>% full_join(pars.ave)
  
  
  env.draw <- rbinom(time.full, 1, env.ratio) # sequence of the environment 
  env.condition <- if_else(env.draw > 0, 'wet','dry')
  
  partitions <- partition_epsilons(sp.inv = sp.2, sp.res = sp.1, pars.full, 
                                   env.condition, time.warm.up, time.full) %>%
    rbind(partition_epsilons(sp.inv = sp.1, sp.res = sp.2, pars.full, 
                             env.condition, time.warm.up, time.full)) %>%
    #pivot_longer(cols = inv.e0:res.eint) %>%
    pivot_longer(cols = inv.full:res.storage) %>%
    separate(name, into = c('player', 'partition')) %>%
    pivot_wider(names_from = player, values_from = value) %>%
    mutate(delta = inv - res,
           partition = factor(partition, 
                              levels = c('full','e0','el','ea','eint', 'storage','nocov')))
  partitions$env.ratio <- env.ratio
  
  partitions.env <- full_join(partitions.env, partitions, 
                              by = c('sp.invader','sp.resident','partition',
                                     'inv','res', 'delta','env.ratio'))
}

partitions.env <- partitions.env  %>%
  mutate(partition = factor(partition, 
                            levels = c('full','e0','el','ea','eint',
                                       'storage','nocov')))

partition.legend <- c(
  'full' = 'Full',  
  'e0' = expression(Delta^0), 
  'el' = expression(Delta^lambda), 
  'ea' = expression(Delta^alpha),
  'eint' = expression(Delta^{alpha*lambda}),
  'storage' = expression(Delta^(alpha*lambda)), 
  'nocov' = expression(Delta^{alpha!=lambda})
)

ggplot(partitions.env, aes(x = env.ratio, y = delta, color = partition)) +
  facet_grid(cols = vars(sp.invader),
             labeller = labeller(sp.invader = species_names)) + 
  geom_line() +
  theme_cw() +
  scale_color_manual(values = cb_palette_long, labels = partition.legend, name = 'Partition') +
  xlab ('Proportion wet years') +
  ylab('Delta')


# Delta as invader-resident comparison
partitions.env.long <- partitions.env %>% 
  pivot_longer(cols = c('delta', 'inv', 'res'), names_to = 'epsilon', values_to = 'value')

ggplot(partitions.env.long, aes(x = env.ratio, y = value, color = partition)) +
  facet_grid(cols = vars(sp.invader), rows = vars(epsilon),
             labeller = labeller(sp.invader = species_names)) + 
  geom_line() +
  theme_cw() +
  scale_color_manual(values = cb_palette_long, labels = partition.legend, name = 'Partition') +
  xlab ('Proportion wet years') +
  ylab('Delta')

```

### Supplemental Figure: simulating coexistence

Verifying that parameters can result in long-term coexistence in conditions predicted by positive low-density growth rates. "invader" and "resident" terms are somewhat arbitrary here as we're starting the species together. But I did also look at this where one species is introduced first to test for priority effects and didn't see any strong differences
```{r long-term-coexistence}

coexistence.env <- tibble(sp.invader = character(), sp.resident = character(), 
                          env.ratio = numeric(), time = numeric(),
                          env.case = character(),
                          an = numeric(), as = numeric(), 
                          s = numeric(), p = numeric())

for(favorable in c('opposite-consistent', 'same-consistent', 'same-flipped')){
  
  for(case in c('p', 'as')){
    pair <- paste(case, 'an', sep = '.')
    
    varying <- 'alpha'
    
    
    # file names
    file.core <- paste0(case, '_', varying, '-', favorable)
    main.wd <- './'
    file.in <- paste0(main.wd, 'model-parameters/model-parameters_', file.core, '.csv')
    #file.out <- paste0(main.wd, 'model-data/mct-partitions_', file.core, '.RData')
    
    # SET PARAMETER VALUES 
    pars <- read.csv(file.in) 
    
    
    # select the relevant parameter set
    if(pair == 'as.an') {
      sp.1 <- 'an'
      sp.2 <- 'as'
      sp.order <- c('an','as')
      pars <- pars %>% mutate(stem.surv = NULL) 
      
    }
    if(pair == 'p.an'){
      sp.1 <- 'an'
      sp.2 <- c('s','p')
      sp.order <- c('an','s','p')
      pars <- pars %>% mutate(seed.surv = NULL)
      
    }
    

    time.full <- 1000
    
    # Starting both species at initially low values
    N0.1 <- 1/length(sp.1)
    N0.2 <- 1/length(sp.2)
    
    env.seq <- seq(0, 1, by = 0.1) # ratio of wet years to dry years
    
    # function to run coexistence invasion for a given env ratio and species,
    # returning average population sizes from the final 1/4 of the time series
    
    coexistence_run <- function(sp.invader, sp.resident, env.ratio, 
                                time.warm.up, time.full, N0.i, N0.r) {
      
      env.draw <- rbinom(time.full, 1, env.ratio) 
      env.condition <- if_else(env.draw > 0, 'wet','dry')
      
      # initializing dataframes
      if(pair == 'as.an'){
        n.all <- tibble(an = as.numeric(rep(NA, time.full)),
                        as = as.numeric(rep(NA, time.full)))
      } else if(pair == 'p.an'){
        n.all <- tibble(an = as.numeric(rep(NA, time.full)),
                        s = as.numeric(rep(NA, time.full)),
                        p = as.numeric(rep(NA, time.full)))
      }
      
      # setting initial populations
      n.all[1,sp.resident] <- N0.r 
      n.all[1,sp.invader] <- N0.i
      
      for(t in 1:time.full){
        # parameters for the varying env condition
        p.env <- pars %>% filter(treatment == env.condition[t])
        
        # running invader and resident together
        N0.together <- n.all[t,]
        
        # running one time step with the appropriate environmental parameters
        n.res.impact <- pop_interactions(N0 = N0.together, param = p.env)
        
        # saving this output as the invader population and the impacted resident population
        n.all[t+1,sp.resident] <- n.res.impact[sp.resident]
        n.all[t+1,sp.invader] <- n.res.impact[sp.invader]
      }
      n.all$time <- 1:(time.full+1)
      
      # average of the last quarter of the time series
      n.ave <- n.all %>% filter(time > time.full*3/4) %>%
        colMeans() %>% as.list()
      n.ave$env.ratio <- env.ratio
      n.ave$sp.invader <- str_c(sp.invader, collapse = ".") # formatting
      n.ave$sp.resident <- str_c(sp.resident, collapse = ".")
      
      return(n.ave)
    }
    
    for(env.ratio in env.seq){
      n.ave.1 <- coexistence_run(sp.invader = sp.1, sp.resident = sp.2, 
                                 env.ratio, time.warm.up, time.full, 
                                 N0.i = N0.1, N0.r = N0.2)
      n.ave.tibble <- n.ave.1 %>% as_tibble() %>% 
        add_column(env.case = favorable)
      coexistence.env <- coexistence.env %>% add_row(n.ave.tibble)
    }
  }
}

#print(coexistence.env)

species.list <- c("an", "as", "s.p")

coexist.env.long <- coexistence.env %>%
  mutate(s.p = p + s, .keep = "unused") %>% 
  pivot_longer(cols = species.list, names_to = "sp", values_to = "population") %>%
  drop_na() 

coexist.env.long$env.case <- coexist.env.long$env.case %>% str_replace_all("-",".")

coexist.env.long$sp.invader <- factor(coexist.env.long$sp.invader, 
                                  levels = c("s.p", "as", "an") )

coexist.env.long$sp.resident <- factor(coexist.env.long$sp.resident, 
                                  levels = c("s.p", "as", "an") )

coexist.env.long$env.case <- factor(coexist.env.long$env.case, 
                                  levels = c("opposite.consistent", "same.flipped", "same.consistent") )

# graph x-axes by proportion of env intrinsically favorable to our focal species
# (perennial or seedbanking annual) -- in the original parameter files
# these were the ones that switched with the different scenarios
coexist.env.long <- coexist.env.long %>% 
  mutate(
    proportion.favorable = ifelse(env.case == "opposite.consistent", 1 - coexist.env.long$env.ratio, coexist.env.long$env.ratio)
  ) 

environmental_condition <- as_labeller(c(
  opposite.consistent = "Opposite env favorable",
  same.consistent = "Same env favorable \n competition stronger",
  same.flipped = "Same env favorable \n competition weaker",
  s.p = "Perennial and \n Nonseedbanking Annual",
  as = "Seedbanking Annual and \n Nonseedbanking Annual"
))

species.legend <- c(
  "an" = "Nonseedbanking Annual",
  "as" = "Seedbanking Annual",
  "s.p" = "Perennial"
)

species.palette <- c("springgreen3", "orangered3", "slateblue4")


ggplot(coexist.env.long, aes(x = proportion.favorable, y = population, color = sp)) +
  facet_grid(rows = vars(env.case), cols = vars(sp.resident), scales = "free",
             labeller = environmental_condition) +
  geom_line(size = 1.1) +
  theme_cw() +
  theme(strip.text = element_text(size = 13),
        legend.position = c(0.17, 0.97),
        legend.text.align = 0,
        legend.text = element_text(size = 11),
        legend.key.size = unit(0.8, "lines")) +
  scale_color_manual(values = species.palette, labels = species.legend, name = '') +
  xlab('Frequency of environment favorable for stage-structured species\' density-independent growth') +
  ylab('Population Size')

 ggsave('mct-partitions_figure-supp_coexist.pdf', width = 8.5, height = 8.5, units = 'in')


```

